### 2021-01-05 Code and miscellaneous/linear-vs-glm-equations.Rmd
Re-organized folder

---
title: "Equations"
author: "John Tipton"
date: "12/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Linear model

$$
y = \beta_0 + time * \beta_1 + treatment * \beta_2 + time * treatment * \beta_3 \\
y = \begin{cases}
\beta_0 + time * \beta_1 & \mbox{if treatment 1}\\
\beta_0 + time * \beta_1 + treatment * \beta_2 + time * treatment * \beta_3 & \mbox{if treatment 2}\\
\end{cases}
\\
$$

plugging in the value treatment = 0 for treatment 1 and the value treatment = 1 for treatment 2

$$
y = \begin{cases}
\beta_0 + time * \beta_1 & \mbox{if treatment 1}\\
( \beta_0 + \beta_2)  + time * (\beta_1 + \beta_3) & \mbox{if treatment 2}\\
\end{cases}
$$

# GLM framework

$$
y = \beta_0 + f(time) * \boldsymbol{\beta}_1 + treatment * \beta_2 + f(time) * treatment * \boldsymbol{\beta}_3 \\
y = \begin{cases}
\beta_0 + f(time) * \boldsymbol{\beta}_1 & \mbox{if treatment 1} \\
\beta_0 + f(time) * \boldsymbol{\beta}_1 + treatment * \beta_2 + f(time) * treatment * \boldsymbol{\beta}_3 & \mbox{if treatment 2}\\
\end{cases}
\\
$$

$$
y = \begin{cases}
\beta_0 + f(time) * \boldsymbol{\beta}_1 & \mbox{if treatment 1}\\
(\beta_0 + \beta_2) + f(time) * (\boldsymbol{\beta}_1 + \boldsymbol{\beta}_3) & \mbox{if treatment 2}\\
\end{cases} \\
= \begin{cases}
\beta_0 + f(time) * \boldsymbol{\beta}_1 & \mbox{if treatment 1}\\
(\tilde{\beta}_2) + f(time) * (\boldsymbol{\tilde{\beta}}_3) & \mbox{if treatment 2}\\
\end{cases}
$$

The Linear model as a GLM: set $f(time) = time$ with the identity function $f(x) = x$