### 2021-01-05 Code and miscellaneous/non-linear-ANOVA.Rmd

---
title: "Non-linear ANOVA example"
output: html_document
---

```{r, message = FALSE}
set.seed(11)
library(patchwork)
library(tidyverse)
```

```{r}
## Example with linear response
example <- function(n_time = 6, 
                    fun_type = "linear", 
                    error_type = "correlated") {

  if (!(fun_type %in% c("linear", "quadratic")))
    stop('fun_type must be either "linear", or "quadratic"')
  if (!(error_type %in% c("correlated", "independent")))
    stop('fun_type must be either "correlated", or "independent"')

  library(tidyverse)
  library(mvnfast)
  library(nlme)

  x <- seq(0, 1, length.out = n_time)
  mu <- matrix(0, length(x), 2)
  # linear response
  if (fun_type == "linear") {
    mu[, 1] <- - (x - .5) + 0.25
    mu[, 2] <- (x - .3)
  } else {
    # nonlinear response
    mu[, 1] <- - 4 * (x - .5)^2 + 0.25
    mu[, 2] <- 4 * (x - .5)^2
  }
  # matplot(mu, type = 'l')

  y <- array(0, dim = c(length(x), 2, 10))
  errors <- array(0, dim = c(length(x), 2, 10))

  if (error_type == "independent") {
    ## independent errors
    for (i in 1:2) {
      for (j in 1:10) {
        errors[, i, j] <- rnorm(6, 0, 0.25)
        y[, i, j] <- mu[, i] + errors[, i, j]
      }
    }
  } else {
    for (i in 1:2) {     # number of treatments
      for (j in 1:10) {  # number of subjects
        # compound symmetry errors
        errors[, i, j] <- rmvn(1, rep(0, length(x)), 0.1 * diag(6) + 0.25 * matrix(1, 6, 6))
        y[, i, j] <- mu[, i] + errors[, i, j]
      }
    }
  }

  ## subject random effects

  ## visualizing the difference between independent errors and compound symmetry
  ## why do we need to account for this -- overly confident inference

  dimnames(y) <- list(time = x, treatment = 1:2, subject = 1:10)
  dimnames(errors) <- list(time = x, treatment = 1:2, subject = 1:10)
  dimnames(mu) <- list(time = x, treatment = 1:2)
  dat <- as.data.frame.table(y, responseName = "y")
  dat_errors <- as.data.frame.table(errors, responseName = "errors")
  dat_mu <- as.data.frame.table(mu, responseName = "mu")
  dat <- left_join(dat, dat_errors, by = c("time", "treatment", "subject"))
  dat <- left_join(dat, dat_mu, by = c("time", "treatment"))
  dat$time <- as.numeric(as.character(dat$time))
  dat <- dat %>%
    mutate(subject = factor(paste(subject, treatment, sep = "-")))

  ## repeated measures ANOVA in R
  fit_lm <- lm(y ~ time + treatment + time * treatment, data = dat)
  dat$preds_lm <- predict(fit_lm)

  fit_lme <- lme(y ~ treatment + time + treatment:time,
                 data = dat,
                 random = ~ 1 | subject,
                 correlation = corCompSymm(form = ~ 1 | subject)
  )

  pred_dat <- expand.grid(
    treatment = factor(1:2), 
    time = unique(dat$time)
  )

  dat$y_pred <- predict(fit_lme)

  return(list(
    dat = dat,
    pred_dat = pred_dat,
    fit_lm = fit_lm,
    fit_lme = fit_lme

  ))
}

plot_example <- function(sim_dat) {
  library(patchwork)
  ## Plot the simulated data
  p1 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = treatment, color = treatment)) +
    geom_point() +
    geom_line(aes(x = time, y = mu, color = treatment)) +
    theme_bw(base_size = 12) +
    ggtitle("Simulated data with true response function")

  p2 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = y, group = subject, color = treatment)) +
    geom_line(aes(size = "Subjects")) +
    # facet_wrap(~ treatment) +
    geom_line(aes(x = time, y = mu, color = treatment, size = "Simulated Truth"), lty = 1) +
    scale_size_manual(name = "Type", values=c("Subjects" = 0.5, "Simulated Truth" = 3)) +
    ggtitle("Simulated data\nIndividual responses with population mean") +
    theme_bw(base_size = 12)

   p3 <- sim_dat$dat %>%
    ggplot(aes(x = time, y = errors, group = subject, color = treatment)) +
    geom_line() +
    # facet_wrap(~ treatment) +
    ggtitle("Simulated errors") +
    theme_bw(base_size = 12)

  p4 <- ggplot(sim_dat$dat, aes(x = time, y = y, color = treatment)) +
    geom_point() +
    geom_line(aes(y = predict(sim_dat$fit_lme), group = subject, size = "Subjects")) +
    geom_line(data = sim_dat$pred_dat, aes(y = predict(sim_dat$fit_lme, level = 0, newdata = sim_dat$pred_dat), size = "Population")) +
    scale_size_manual(name = "Predictions", values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_bw(base_size = 12) +
    ggtitle("Fitted Model")
  return((p1 + p3) / (p2 + p4))
  # return((p1 + p3 + plot_layout(guides = "collect")) / (p2 + p4 + plot_layout(guides = "collect")))

}
```

```{r, fig.width=16, fig.height=9, out.width="100%"}
set.seed(111)
# linear response, independent errors (no subject effect)
plot_example(example(fun_type = "linear", error_type = "independent")) +
  plot_annotation(
  title = 'Simulated with linear function and independent errors', 
  theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```

```{r, fig.width=16, fig.height=9, out.width="100%"}
# linear response, correlated errors (subject effect)
plot_example(example(fun_type = "linear", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with linear function and correlated errors',
    theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```

```{r, fig.width=16, fig.height=9, out.width="100%"}
# quadratic response, independent errors (no subject effect)
plot_example(example(fun_type = "quadratic", error_type = "independent")) +
  plot_annotation(
  title = 'Simulated with quadratic function and independent errors',
    theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```

```{r, fig.width=16, fig.height=9, out.width="100%"}
# quadratic response, correlated errors (subject effect)
plot_example(example(fun_type = "quadratic", error_type = "correlated")) +
  plot_annotation(
  title = 'Simulated with quadratic function and correlated errors',
    theme = theme(plot.title = element_text(size = 20))
) #& 
  # theme(plot.title = element_text(size = 20))
```