### 2021-05-02 Code and miscellaneous/Tumor response GAM with comments.Rmd
Created RMarkdown document

---
title: "Tumor_Response_GAM_comments"
author: "Ariel Mundo"
date: "5/2/2021"
output: html_document
---

```{r,loading packages}
library(mgcv)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(gratia)
```

number of observations -- currently using 5
use measurement error of sd as 10 (rough estimate)

Data are simulated from Fig. 3 (C) of Vishwanath et. al:
Vishwanath, Karthik, et al. "Using optical spectroscopy to
longitudinally monitor physiological changes within solid tumors."
Neoplasia 11.9 (2009): 889-900.
https://www.sciencedirect.com/science/article/pii/S1476558609800420

```{r,simulating data}
dat <- read_csv(here::here("data", "Vishwanath StO2.csv"))
dat$Group<-as.factor(dat$Group)
## plot the mean response
ggplot(dat, aes(x = Day, y = StO2, color = Group)) +
    geom_line()+
    geom_point()+
    scale_x_continuous(breaks=c(0,2,5,7,10))

#This function simulates data for the tumor data using a dafault number of
#10 observations per time point,
#and a standard deviation of 10% StO2
#Because physiologically StO2 cannot go below 0%, data is generated with a cutoff value of
#0.0001 (the "observation_tobit")

simulate_data <- function(dat, n = 10, sd = 5) {
    dat_sim <- dat %>%
        slice(rep(1:n(), each = n)) %>%
        group_by(Group, Day) %>%
        mutate(observation = rnorm(n, StO2, sd),
               observation_tobit = pmax(rnorm(n, StO2, sd), 0.0001)) %>%
        ungroup()

    return(dat_sim)
}

n <- 10
sd <- 10
set.seed(11) #setting seed
df <- 6
dat_sim <- simulate_data(dat, n, sd)

#plotting simulated data
ggplot(dat_sim, aes(x = Day, y = observation_tobit, color = Group)) +
    geom_point()+
    scale_x_continuous(breaks=c(0,2,5,7,10))

n_time <- length(unique(dat_sim$Day))
n_treatment <- length(unique(dat_sim$Group))

#if you have n timepoints can you do n knots or should you do n-1 knots? (I read somewhere that
#the n-1 option was preferred for computational issues)?

```

```{r,models}
mod_simple  <- gam(observation_tobit ~ s(Day,by=Group,k=5),data  = dat_sim)
appraise(mod2)
#SIMPLEST MODEL: A SMOOTH FOR TIME AS THE COVARIATE, WHICH CAN VARY BY GROUP BUT... APPRAISING THE MODEL SHOWS ISSUES ON THE RESIDUALS AND THE DATA DISTRIBUTION

#Model with interaction using tensor
mod_int1 <- gam(observation_tobit ~ Group+te(Day, by = Group, k = 5), data  = dat_sim)
appraise(mod1)

#Model with interaction using s
mod3 <- gam(observation_tobit ~ Group+s(Day, by = Group, k = 5), data  = dat_sim)
appraise(mod3)
```

Ok, so mod1 and mod3 are doing essentially the same thing. We have a smooth of time
per group, and we take into account the main effect of each treatment.
So this model effectively takes into account treatment and time? and the simples model only takes the effect of time.

"To avoid confounding problems the smooths are all subject to sum to zero constraints, which usually means that the main effect of g should also be included
in the model specification. For example, g + te(z,x,by=g)(pg.327)" (From Simon Wood's book)

The idea for this came from
https://stats.stackexchange.com/questions/486118/model-building-in-generalized-additive-mixed-models-gamms

```{r, plotting model}
#plotting smooths and simulated data
#full data
#I am doing something wrong here but I can't figure it out!!!!
#f_predict<-with(dat_sim,expand.grid(StO2=seq(min(observation_tobit),
#                                                        max(observation_tobit),
#                                                         length=10),
#                                                     Group=Group,Day=Day))

 #f_predict<-cbind(f_predict,
   #                       predict(mod2,f_predict,
  #                                se.fit = TRUE,type='response'))
#

#creates a dataframe wusing the lenght of the covariates
f_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1))

#adds the predictions to the grid and creates a confidence interval
f_predict<-f_predict%>%
    mutate(fit = predict(mod1,f_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(mod1, f_predict,se.fit = TRUE,type='response')$se.fit)

ggplot(data=dat_sim, aes(x=Day, y=observation_tobit, group=Group)) +
    facet_wrap(~Group)+
    geom_point(colour='black',size=1,alpha=0.5)+
    geom_ribbon(aes(ymin=(fit - 2*se.fit), ymax=(fit + 2*se.fit), x=Day),
                data=f_predict, alpha=0.5,
                inherit.aes=FALSE) +
    geom_line(aes(y=(fit),color=factor(Group)),
              size=1,data=f_predict,show.legend = FALSE)

#it works now, but why do the smooths still look so rough? (the connections at the knots
#don't look "smooth")

##trying the pairwise comparisons

pdat <- expand.grid(Day = seq(0, 10, length = 400),
                    Group = c('Treatment', 'Control'))
smooth_diff <- function(model, newdata, f1, f2, alpha = 0.05,
                        unconditional = FALSE) {
    xp <- predict(model, newdata = newdata, type = 'lpmatrix')
    c1 <- grepl(f1, colnames(xp))
    c2 <- grepl(f2, colnames(xp))
    #r1 <- newdata[[var]] == f1
    #r2 <- newdata[[var]] == f2
    r1 <- with(newdata, Group == f1)
    r2 <- with(newdata, Group == f2)
    ## difference rows of xp for data from comparison
    X <- xp[r1, ] - xp[r2, ]
    ## zero out cols of X related to splines for other lochs
    X[, ! (c1 | c2)] <- 0
    ## zero out the parametric cols
    X[, !grepl('^s\\(', colnames(xp))] <- 0
    dif <- X %*% coef(model)
    se <- sqrt(rowSums((X %*% vcov(model, unconditional = unconditional)) * X))
    crit <- qt(alpha/2, df.residual(model), lower.tail = FALSE)
    upr <- dif + (crit * se)
    lwr <- dif - (crit * se)
    data.frame(pair = paste(f1, f2, sep = '-'),
               diff = dif,
               se = se,
               upper = upr,
               lower = lwr)
}
comp1<-smooth_diff(mod2,pdat,'Control','Treatment')
comp_StO2 <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp1))

c1<-ggplot(comp_StO2, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))

c1
    ##################################################################
############################################################
#missing data
#create a sequence of 40 random numbers between 1 and 100, these numbers will
#correspond to the row numbers to be randomly erased from the original dataset
missing <- sample(1:100, 40)

#create a new dataframe from the simulated data with 40 rows randomly removed
dat_missing <- dat_sim[-missing, ]

#Count the number of remaining observations per day (original dataset had 10 per group per day)
dat_missing %>%
    group_by(Day, Group) %>%
    summarize(count = n())

#plot the simulated data
ggplot(dat_missing, aes(x = Day, y = observation_tobit, color = Group)) +
    geom_point() +
    ggtitle("Simulated time-varying responses") +
    scale_color_viridis_d(end = 0.75) +
    geom_line(aes(y = StO2), alpha = 0.75) +
    theme_bw()

#the same model used for the full dataset
mod_m1 <- gam(observation_tobit ~ Group+s(Day,by=Group,k=5), data  = dat_missing)
#appraise the model
appraise(mod_m1)

m_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1))

#adds the predictions to the grid and creates a confidence interval
m_predict<-m_predict%>%
    mutate(fit = predict(mod_m1,m_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(mod_m1, m_predict,se.fit = TRUE,type='response')$se.fit)

ggplot(data=dat_missing, aes(x=Day, y=observation_tobit, group=Group)) +
    facet_wrap(~Group)+
    geom_point(colour='black',size=1,alpha=0.5)+
    geom_ribbon(aes(ymin=(fit - 2*se.fit), ymax=(fit + 2*se.fit), x=Day),
                data=m_predict, alpha=0.5,
                inherit.aes=FALSE) +
    geom_line(aes(y=(fit),color=factor(Group)),
              size=1,data=m_predict,show.legend = FALSE)

comp2<-smooth_diff(mod_m1,pdat,'Control','Treatment')
comp_StO2_m <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp2))

c2<-ggplot(comp_StO2_m, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    labs(x = 'Days', y = expression(paste('Difference in StO'[2],'(missing data)' )))

c2

c1+c2
```