### 2021-05-04 Code and miscellaneous/Tumor_response_GAM__comments.Rmd
Compiled pdf for RMarkdown document of GAM model for tumor data

---
title: "Tumor Response GAM models with comments"
author: "Ariel Mundo"
date: "5/2/2021"
output:
  pdf_document: default
  html_document: default
---

```{r,loading packages,echo=FALSE,include=FALSE,message=FALSE,warning=FALSE}
library(mgcv)
library(tidyverse)
library(ggplot2)
library(patchwork)
library(gratia)
library (ggsci)
```

number of observations -- currently using 5
use measurement error of sd as 10 (rough estimate)

Data are simulated from Fig. 3 (C) of Vishwanath et. al:
"Using optical spectroscopy to longitudinally monitor physiological changes within solid tumors." Neoplasia 11.9 (2009): 889-900.
https://www.sciencedirect.com/science/article/pii/S1476558609800420

From the figure the maximum standard deviation for the measurements is approximately 10%.

```{r,simulating data,cho=FALSE,include=FALSE,message=FALSE,warning=FALSE}
dat <- read_csv(here::here("data", "Vishwanath StO2.csv"))
dat$Group<-as.factor(dat$Group)

## plot the mean response
f1<-ggplot(dat, aes(x = Day, y = StO2, color = Group)) +
    geom_line(size=1,show.legend = FALSE)+
    geom_point(show.legend = FALSE)+
  labs(y=expression(paste('Mean StO'[2],' (real)')))+
  theme_classic()+
  scale_color_aaas()+
    scale_x_continuous(breaks=c(0,2,5,7,10))

#This function simulates data for the tumor data using default parameters of 10 observations per time point,and Standard deviation (sd) of 5%.
#Because physiologically StO2 cannot go below 0%, data is  generated with a cutoff value of 0.0001 (the "StO2_sim")

simulate_data <- function(dat, n = 10, sd = 5) {
    dat_sim <- dat %>%
        slice(rep(1:n(), each = n)) %>%
        group_by(Group, Day) %>%
        mutate(
               StO2_sim = pmax(rnorm(n, StO2, sd), 0.0001)) %>%
        ungroup()

    return(dat_sim)
}

n <- 10 #number of observations
sd <- 10 #approximate sd from paper
set.seed(1) #set seed for reproducibility
df <- 6
dat_sim <- simulate_data(dat, n, sd)

#plotting simulated data
f2<-ggplot(dat_sim, aes(x = Day, y = StO2_sim, color = Group)) +
    geom_point()+
    stat_summary(aes(y = StO2_sim,
                     group=Group), 
                 fun=mean, geom="line",size=1)+
  labs(y=expression(paste('StO'[2],' (simulated)')))+
  theme_classic()+
  scale_color_aaas()+
    scale_x_continuous(breaks=c(0,2,5,7,10))

#plot the "true" data with the simulated response

f1+f2+plot_layout(guides = 'collect') &
  ylim(c(0,75)) &
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )&
  scale_color_aaas()

```

```{r,models}
mod_simple  <- gam(StO2_sim ~ s(Day,by=Group,k=5),data  = dat_sim)
appraise(mod_simple)
#SIMPLEST MODEL: A SMOOTH FOR TIME AS THE COVARIATE, WHICH CAN VARY BY GROUP BUT... APPRAISING THE MODEL SHOWS ISSUES ON THE RESIDUALS AND THE DATA DISTRIBUTION

#Model with interaction using tensor: there is some different specification that does not allow to plot it easily
mod_int1 <- gam(StO2_sim ~ Group+te(Day,by=Group, k = 5),
                data  = dat_sim)
appraise(mod_int1)

#Model with interaction using s
mod_int2 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5),method='REML', data  = dat_sim)
appraise(mod_int2)
```

The simplest model attempts to explain how $StO_2$ changes over time, and gives each treatment group a separate smooth. However, from the diagnostics, it is clear that there are some issues because the QQ plot, the residuals vs linear predictor and the histogram show that the assumptions of normality are not met.

Next, we build a model that accounts for the _interaction_ of time and treatment. In this sense, mod_int1 and mod_int2 are doing essentially the same thing: We have a smooth to explain how $StO_2$ changes per group, but we also account separately for the "main effect" of treatment. In other words, we are trying to explain the changes in $StO_2$ by considering how it changes over time, but also disentangling the treatment effect from the response.

"To avoid confounding problems the smooths are all subject to sum to zero constraints, which usually means that the main effect of g should also be included
in the model specification. For example, g + te(z,x,by=g)(pg.327)" (From Simon Wood's book). This gives a good explanation on why mod_int1 and mod_int2 work.

The idea for this came from
https://stats.stackexchange.com/questions/486118/model-building-in-generalized-additive-mixed-models-gamms

```{r, plotting model,echo=FALSE,include=FALSE,message=FALSE,warning=FALSE}
#plotting smooths and simulated data

#creates a dataframe using the length of the covariates
f_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1))

#adds the predictions to the grid and creates a confidence interval
f_predict<-f_predict%>%
    mutate(fit = predict(mod_int2,f_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(mod_int2, f_predict,se.fit = TRUE,type='response')$se.fit)

#plot smooths and confidence interval
f3<-ggplot(data=dat_sim, aes(x=Day, y=StO2_sim, group=Group)) +
    geom_point(aes(color=Group),size=1.5,alpha=0.5)+
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),fill=Group),
              alpha=0.3,
              data=f_predict,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=f_predict,
              show.legend = FALSE)+
  facet_wrap(~Group)+
  labs(y=expression(paste('StO'[2],' (simulated)')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
      scale_color_aaas()+
  scale_fill_aaas()

f3 & theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

##Pairwise comparisons

pdat <- expand.grid(Day = seq(0, 10, length = 400),
                    Group = c('Treatment', 'Control'))

smooth_diff <- function(model, newdata, f1, f2, alpha = 0.05,
                        unconditional = FALSE) {
    xp <- predict(model, newdata = newdata, type = 'lpmatrix')
    c1 <- grepl(f1, colnames(xp))
    c2 <- grepl(f2, colnames(xp))
    #r1 <- newdata[[var]] == f1
    #r2 <- newdata[[var]] == f2
    r1 <- with(newdata, Group == f1)
    r2 <- with(newdata, Group == f2)
    ## difference rows of xp for data from comparison
    X <- xp[r1, ] - xp[r2, ]
    ## zero out cols of X related to splines for other lochs
    X[, ! (c1 | c2)] <- 0
    ## zero out the parametric cols
    X[, !grepl('^s\\(', colnames(xp))] <- 0
    dif <- X %*% coef(model)
    se <- sqrt(rowSums((X %*% vcov(model, unconditional = unconditional)) * X))
    crit <- qt(alpha/2, df.residual(model), lower.tail = FALSE)
    upr <- dif + (crit * se)
    lwr <- dif - (crit * se)
    data.frame(pair = paste(f1, f2, sep = '-'),
               diff = dif,
               se = se,
               upper = upr,
               lower = lwr)
}

comp1<-smooth_diff(mod_int2,pdat,'Control','Treatment')

comp_StO2 <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp1))

c1<-ggplot(comp_StO2, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.5,fill='#993336') +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
  theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
  scale_x_continuous(breaks=c(0,2,5,7,10))+
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

c1
```

GAMs do not require equally-spaced observations and can accommodate missing observations as well. In a case where 40% of the observations are missing:

```{r,missing data,echo=FALSE,include=TRUE,message=FALSE,warning=FALSE}
    ##################################################################
############################################################
#missing data
#create a sequence of 40 random numbers between 1 and 100, these numbers will
#correspond to the row numbers to be randomly erased from the original dataset
missing <- sample(1:100, 40)

#create a new dataframe from the simulated data with 40 rows randomly removed, keep the missing values as NA

ind <- which(dat_sim$StO2_sim %in% sample(dat_sim$StO2_sim, 40))

#create a new dataframe, remove the StO2 column
dat_missing <- dat_sim[,-1]

#add NAs at the ind positions
dat_missing$StO2_sim[ind]<-NA

#Count the number of remaining observations per day (original dataset had 10 per group per day)
dat_missing %>%
    group_by(Day,Group) %>%
    filter(!is.na(StO2_sim))%>%
  count(Day)

#plot the simulated data
f4<-ggplot(dat_missing, aes(x = Day, y = StO2_sim, color = Group)) +
    geom_point() +
  stat_summary(aes(y = StO2_sim,
                     group=Group), 
                 fun=mean, geom="line",size=1)+
    ggtitle("Missing data") +
  labs(y='StO2')+
  scale_color_aaas()+
    theme_classic()

f2+f4+plot_layout(guides = 'collect') &
  ylim(c(0,75)) &
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )&
  scale_color_aaas()

#the same model used for the full dataset
mod_m1 <- gam(StO2_sim ~ Group+s(Day,by=Group,k=5), data  = dat_missing)
#appraise the model
appraise(mod_m1)

m_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1))

#adds the predictions to the grid and creates a confidence interval
m_predict<-m_predict%>%
    mutate(fit = predict(mod_m1,m_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(mod_m1, m_predict,se.fit = TRUE,type='response')$se.fit)

ggplot(data=dat_missing, aes(x=Day, y=StO2_sim, group=Group)) +
    facet_wrap(~Group)+
    geom_point(colour='black',size=1,alpha=0.5)+
    geom_ribbon(aes(ymin=(fit - 2*se.fit), ymax=(fit + 2*se.fit), x=Day),
                data=m_predict, alpha=0.5,
                inherit.aes=FALSE) +
    geom_line(aes(y=(fit),color=factor(Group)),
              size=1,data=m_predict,show.legend = FALSE)

comp2<-smooth_diff(mod_m1,pdat,'Control','Treatment')
comp_StO2_m <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp2))

c2<-ggplot(comp_StO2_m, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    labs(x = 'Days', 
         y = expression(paste('Difference in StO'[2],'\n (missing data)' 
                              )))+
  scale_x_continuous(breaks=c(0,2,5,7,10))+
  theme_classic()+
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

c2

c1+c2
```

It is reassuring to see that we get the same diagnostics by either having a dataframe with missing observations as NA or a dataframe where the missing observations are not present at all.

So two questions remains after this:
-How do we want to present the pairwise smooth confidence interval comparison? 
The GAMs are referred as "empirical Bayesian", but I could not find a straightforward reference that had information in detail of what the pairwise comparisons can tell at certain timepoints. Sure, they are different, but what else?

-Do we want to address prediction? I recently saw a webinar from Gavin Simpson regarding this, wondered if it would be feasible to include in a concise manner.