### 2021-06-01 Manuscripts/Manuscript by chapters -v2/06-Conclusion.Rmd

# Conclusion

We have presented GAMs as a method to analyze longitudinal biomedical data. Future directions of this work will include simulation-based estimations of statistical power using GAMs, as well as demonstrating the prediction capabilities of these models using large datasets. 
By making the data and code used in this paper accessible, we hope to address the need of creating and sharing reproducible work in biomedical research.

# Acknowledgements

This work was supported by the National Science Foundation Career Award (CBET 1751554) and the Arkansas Biosciences Institute.