### 2021-06-18 Manuscripts/Manuscript by chapters -for submission/Full_document_SAGE.tex

\documentclass[Royal,times,sagev]{sagej}

\usepackage{moreverb,url,natbib, multirow, tabularx}
\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

% Pandoc citation processing

\begin{document}

\title{Using generalized additive models to analyze biomedical
non-linear longitudinal data}

\runninghead{Mundo \emph{et al}.}

\author{Ariel I. Mundo\affilnum{1}, John R. Tipton\affilnum{2}, Timothy
J. Muldoon*\affilnum{1}}

\affiliation{\affilnum{1}{Department of Biomedical Engineering,
University of Arkansas, Fayetteville, AR, US}\\\affilnum{2}{Department
of Mathematical Sciences, University of Arkansas, Fayetteville, AR, US}}

\corrauth{Corresponding author}

\email{\href{mailto:tmuldoon@uark.edu}{\nolinkurl{tmuldoon@uark.edu}}}

\begin{abstract}
In biomedical research, the outcome of longitudinal studies has been
traditionally analyzed using the \emph{repeated measures analysis of
variance} (rm-ANOVA) or more recently, \emph{linear mixed models}
(LMEMs). Although LMEMs are less restrictive than rm-ANOVA in terms of
correlation and missing observations, both methodologies share an
assumption of linearity in the measured response, which results in
biased estimates and unreliable inference when they are used to analyze
data where the trends are non-linear, which is a common occurrence in
biomedical research. In contrast, \emph{generalized additive models}
(GAMs) relax the linearity assumption, and allow the data to determine
the fit of the model while permitting missing observations and different
correlation structures. Therefore, GAMs present an excellent choice to
analyze non-linear longitudinal data in the context of biomedical
research. This paper summarizes the limitations of rm-ANOVA and LMEMs
and uses simulated data to visually show how both methods produce biased
estimates when used on non-linear data. We also present the basic theory
of GAMs, and using trends of oxygen saturation in tumors reported in the
biomedical literature, we simulate example longitudinal data (2
treatment groups, 10 subjects per group, 5 repeated measures for each
group) to demonstrate how these models can be computationally
implemented. We show that GAMs are able to produce estimates that are
consistent with the trends of biomedical non-linear data even in the
case when missing observations exist (with 40\% of the simulated
observations missing), allowing reliable inference from the data. To
make this work reproducible, the code and data used in this paper are
available at: \url{https://github.com/aimundo/GAMs-biomedical-research}.
\end{abstract}

\keywords{Cancer biology; Tumor response; generalized additive models;
simulation; R}

\maketitle

\hypertarget{background}{%
\section{Background}\label{background}}

Longitudinal studies are designed to repeatedly measure a variable of
interest in a group (or groups) of subjects, with the intention of
observing the evolution of effect across time rather than analyzing a
single time point (e.g., a cross-sectional study). Biomedical research
frequently uses longitudinal studies to analyze the evolution of a
``treatment'' effect across multiple time points; and in such studies
the subjects of analysis range from animals (mice, rats, rabbits), to
human patients, cells, or blood samples, among many others. Tumor
response \citep{roblyer2011, tank2020, pavlov2018, demidov2018},
antibody expression \citep{ritter2001, roth2017}, and cell metabolism
\citep{jones2018, skala2010} are examples of the different situations
where researchers have used longitudinal designs to study some
physiological response. Because the frequency of the measurements in a
longitudinal study is dependent on the biological phenomena of interest
and the experimental design of the study, the frequency of such
measurements can range from minute intervals to study a short-term
response such as anesthesia effects in animals\citep{greening2018}, to
weekly measurements to analyze a mid-term response like the evolution of
dermatitis symptoms in breast cancer patients \citep{sio2016}, to
monthly measurements to study a long-term response such as mouth opening
following radiotherapy (RT) in neck cancer patients \citep{kamstra2015}.

Traditionally, a ``frequentist'' or ``classical'' statistical paradigm
is used in biomedical research to derive inferences from a longitudinal
study. The frequentist paradigm regards probability as the limit of the
expected outcome when an experiment is repeated a large number of times
\citep{wagenmakers2008}, and such view is applied to the analysis of
longitudinal data by assuming a null hypothesis under a statistical
model that is often an \emph{analysis of variance over repeated
measures} (repeated measures ANOVA or rm-ANOVA). The rm-ANOVA model
makes three key assumptions regarding longitudinal data: 1) linearity of
the response across time, 2) constant correlation across same-subject
measurements, and 3) observations from each subject are obtained at all
time points through the study (a condition also known as \emph{complete
observations}) \citep{gueorguieva2004, schober2018}.

The expected linear behavior of the response through time is a key
requisite in rm-ANOVA \citep{pinheiro2006}. This ``linearity
assumption'' in rm-ANOVA implies that the model is misspecified when the
data does not follow a linear trend, which results in unreliable
inference. In biomedical research, non-linear trends are the norm rather
than the exception in longitudinal studies. A particular example of this
non-linear behavior in longitudinal data arises in measurements of tumor
response to chemo and/or radiotherapy in preclinical and clinical
settings \citep{roblyer2011, skala2010, vishwanath2009}. These studies
have shown that the collected signal does not follow a linear trend over
time, and presents extreme variability at different time points, making
the fit of rm-ANOVA model inconsistent with the observed variation.
Therefore, when rm-ANOVA is used to draw inference of such data the
estimates are inevitably biased, because the model is only able to
accommodate linear trends that fail to adequately represent the
biological phenomenon of interest.

A \emph{post hoc} analysis is often used in conjunction with rm-ANOVA to
perform repeated comparisons to estimate a \emph{p-value}, which in turn
is used as a measure of significance. Although it is possible that a
\emph{post hoc} analysis of rm-ANOVA is able to find ``significant''
\emph{p-values}( \emph{p}\textless0.05) from non-linear data, the
validity of such metric is dependent on how adequate the model fits the
data. In other words, \emph{p-values} are valid only if the model and
the data have good agreement; if that is not the case, a ``Type III''
error (known as ``model misspecification'') occurs\citep{dennis2019}.
For example, model misspecification will occur when a model that is only
able to explain linear responses (such as rm-ANOVA) is fitted to data
that follows a quadratic trend, thereby causing the resulting
\emph{p-values} and parameter estimates to be invalid \citep{wang2019}.

Additionally, the \emph{p-value} itself is highly variable, and multiple
comparisons can inflate the false positivity rate (Type I error or
\(\alpha\)) \citep{liu2010, halsey2015}, consequently biasing the
conclusions of the study. Corrections exist to address the Type I error
issue of multiple comparisons (such as Bonferroni \citep{abdi2010}), but
they in turn reduce statistical power (1-\(\beta\))\citep{nakagawa2004},
and lead to increased Type II error (failing to reject the null
hypothesis when the null hypothesis is false)
\citep{gelman2012, albers2019}. Therefore, the tradeoff of \emph{post
hoc} comparisons in rm-ANOVA between Type I, II and III errors might be
difficult to resolve in a biomedical longitudinal study where a delicate
balance exists between statistical power and sample size.

On the other hand, the assumption of constant correlation in rm-ANOVA
(often known as the \emph{compound symmetry assumption}) is typically
unreasonable because correlation between the measured responses often
diminishes as the time interval between the observation increases
\citep{ugrinowitsch2004}. Corrections can be made in rm-ANOVA in the
absence of compound symmetry \citep{huynh1976, greenhouse1959}, but the
effectiveness of the correction is limited by the size of the sample,
the number of measurements\citep{haverkamp2017}, and group sizes
\citep{keselman2001}. In the case of biomedical research, where living
subjects are frequently used, sample sizes are often not ``large'' due
to ethical and budgetary reasons \citep{charan2013} which might cause
the corrections for lack of compound symmetry to be ineffective.

Due to a variety of causes, the number of observations during a study
can vary between all subjects. For example, in a clinical trial patients
may voluntarily withdraw, whereas attrition due to injury or weight loss
in preclinical animal studies is possible. It is even plausible that
unexpected complications with equipment or supplies arise that prevent
the researcher from collecting measurements at certain time points. In
each of these missing data scenarios, the \emph{complete observations}
assumption of classical rm-ANOVA is violated. When incomplete
observations occur, a rm-ANOVA model is fit by excluding all subjects
with missing observations from the analysis \citep{gueorguieva2004}.
This elimination of partially missing data from the analysis can result
in increased costs if the desired statistical power is not met with the
remaining observations, because it would be necessary to enroll more
subjects. At the same time, if the excluded observations contain
insightful information that is not used, their elimination from the
analysis may limit the demonstration of significant differences between
groups.

During the last decade, the biomedical community has started to
recognize the limitations of rm-ANOVA in the analysis of longitudinal
data. The recognition on the shortcomings of rm-ANOVA is exemplified by
the use of linear mixed effects models (LMEMs) by certain groups to
analyze longitudinal tumor response data
\citep{skala2010, vishwanath2009}. Briefly, LMEMs incorporate
\emph{fixed effects}, which correspond to the levels of experimental
factors in the study (e.g., the different drug regimens in a clinical
trial), and \emph{random effects}, which account for random variation
within the population (e.g., the individual-level differences not due to
treatment such as weight or age). When compared to the traditional
rm-ANOVA, LMEMs are more flexible as they can accommodate missing
observations for multiple subjects and allow different modeling
strategies for the variability within each measure in every subject
\citep{pinheiro2006}. However, LMEMs impose restrictions in the
distribution of the errors of the random effects, which need to be
normally distributed and independent \citep{gueorguieva2004, barr2013}.
And even more importantly, LMEMs also assume a linear relationship
between the response and time \citep{pinheiro2006}, making them
unsuitable to analyze non-linear data.

As the rm-ANOVA and the more flexible LMEM approaches make overly
restrictive assumptions regarding the linearity of the response, there
is a need for biomedical researchers to explore the use of additional
statistical tools that allow the data (and not an assumption in trend)
to determine the trend of the fitted model, to enable appropriate
inference. In this regard, generalized additive models (GAMs) present an
alternative approach to analyze longitudinal data. Although not
frequently used by the biomedical community, these semi-parametric
models are customarily used in other fields to analyze longitudinal
data. Examples of the use of GAMs include the analysis of temporal
variations in geochemical and palaeoecological data
\citep{rose2012, pedersen2019, simpson2018}, health-environment
interactions \citep{yang2012} and the dynamics of government in
political science \citep{beck1998} . There are several advantages of
GAMs over LMEMs and rm-ANOVA models: 1) GAMs can fit a more flexible
class of smooth responses that enable the data to dictate the trend in
the fit of the model, 2) they can model non-constant correlation between
repeated measurements \citep{wood2017} and 3) can easily accommodate
missing observations. Therefore, GAMs can provide a more flexible
statistical approach to analyze non-linear biomedical longitudinal data
than LMEMs and rm-ANOVA.

The current advances in programming languages designed for statistical
analysis (specifically \(\textsf{R}\)), have eased the computational
implementation of traditional models such as rm-ANOVA and more complex
approaches such as LMEMs and GAMs. In particular, \(\textsf{R}\)
\citep{r} has an extensive collection of documentation and functions to
fit GAMs in the package \emph{mgcv} \citep{wood2016, wood2017} that not
only speed up the initial stages of the analysis but also enable the use
of advanced modeling structures (e.g.~hierarchical models, confidence
interval comparisons) without requiring advanced programming skills from
the user. At the same time, \(\textsf{R}\) has many tools that simplify
data simulation, an emerging strategy used to test statistical models
\citep{haverkamp2017}. Data simulation methods allow the researcher to
create and explore different alternatives for analysis without
collecting information in the field, reducing the time window between
experiment design and its implementation, and simulation can be also
used for power calculations and study design questions.

This work provides biomedical researchers with a clear understanding of
the theory and the practice of using GAMs to analyze longitudinal data
using by focusing on four areas. First, the limitations of LMEMs and
rm-ANOVA regarding linearity of response, constant correlation
structures and missing observations are explained in detail. Second, the
key theoretical elements of GAMs are presented using clear and simple
mathematical notation while explaining the context and interpretation of
the equations. Third, we illustrate the type of non-linear longitudinal
data that often occurs in biomedical research using simulated data that
reproduces patterns in previously reported studies
\citep{vishwanath2009}. The simulated data experiments highlight the
differences in inference between rm-ANOVA, LMEMs and GAMs on data
similar to what is commonly observed in biomedical studies. Finally,
reproducibility is emphasized by providing the code to generate the
simulated data and the implementation of different models in
\(\textsf{R}\), in conjunction with a step-by-step guide demonstrating
how to fit models of increasing complexity.

In summary, this work will allow biomedical researchers to identify when
the use of GAMs instead of rm-ANOVA or LMEMs is appropriate to analyze
longitudinal data, and provide guidance on the implementation of these
models to improve the standards for reproducibility in biomedical
research.

\hypertarget{challenges-presented-by-longitudinal-studies}{%
\section{Challenges presented by longitudinal
studies}\label{challenges-presented-by-longitudinal-studies}}

\hypertarget{the-repeated-measures-anova-and-linear-mixed-model}{%
\subsection{The repeated measures ANOVA and Linear Mixed
Model}\label{the-repeated-measures-anova-and-linear-mixed-model}}

The \emph{repeated measures analysis of variance} (rm-ANOVA) and the
\emph{linear mixed model} (LMEM) are the most commonly used statistical
analysis for longitudinal data in biomedical research. These statistical
methodologies require certain assumptions for the model to be valid.
From a practical view, the assumptions can be divided in three areas: 1)
linear relationship between covariates and response, 2) a constant
correlation between measurements, and, 3) complete observations for all
subjects. Each one of these assumptions is discussed below.

\hypertarget{linear-relationship}{%
\subsection{Linear relationship}\label{linear-relationship}}

\hypertarget{the-repeated-measures-anova-case}{%
\subsubsection{The repeated measures ANOVA
case}\label{the-repeated-measures-anova-case}}

In a longitudinal biomedical study, two or more groups of subjects
(e.g., human subject, mice, samples) are subject to different treatments
(e.g., a ``treatment'' group receives a novel drug or intervention vs.~a
``control'' group that receives a placebo), and measurements from each
subject within each group are collected at specific time points. The
collected response is modeled with \emph{fixed} components. The
\emph{fixed} component can be understood as a constant value in the
response which the researcher is interested in measuring, i.e., the
average effect of the novel drug/intervention in the ``treatment''
group.

Mathematically speaking, a rm-ANOVA model with an interaction can be
written as:

\begin{equation}
y_{ijt} = \beta_0+\beta_1 \times time_{t} +\beta_2 \times treatment_{j} +\beta_3 \times time_{t}\times treatment_{j}+\varepsilon_{ijt}\\ 
(\#eq:linear-model)
\end{equation}

In this model \(y_{ijt}\) is the response for subject \(i\), in
treatment group \(j\) at time \(t\), which can be decomposed in a mean
value \(\beta_0\), \emph{fixed effects} of time (\(time_t\)), treatment
(\(treatment_j\)) and their interaction \(time_t*treatment_j\) which
have linear slopes given by \(\beta_1, \beta_2\) and \(\beta_3\),
respectively. Independent errors \(\varepsilon_{ijt}\) represent random
variation not explained by the \emph{fixed} effects, and are assumed to
be \(\sim N(0,\sigma^2)\) (independently and identically normally
distributed with mean zero and variance \(\sigma^2\)). In a biomedical
research context, suppose two treatments groups are used in a study
(e.g., ``placebo'' vs.~``novel drug'' or ``saline''
vs.~``chemotherapy''). Then, the group terms in Equation
@ref(eq:linear-model) can be written as below with \(treatment_j=0\)
representing the first treatment group (Group A) and \(treatment_j=1\)
representing the second treatment group (Group B). With this notation,
the linear model then can be expressed as

\begin{equation}
y_{ijt} = \begin{cases}
\beta_0 + \beta_1\times time_{t}+\varepsilon_{ijt}   & \mbox{if Group A}\\
\beta_0 + \beta_2+\beta_1 \times time_{t} +\beta_3 \times time_{t}+\varepsilon_{ijt}  & \mbox{if Group B}\\
\end{cases}
(\#eq:ANOVA-by-group)
\end{equation}

To further simplify the expression, substitute
\(\widetilde{\beta_{0}}=\beta_0+\beta_{2}\) and
\(\widetilde{\beta_{1}}=\beta_{1}+\beta_{3}\) in the equation for Group
B. This substitution allows for a different intercept and slope for
Groups A and B. The model is then written as

\begin{equation}
y_{ijt} = \begin{cases}
\beta_0 + \beta_1\times time_{t}+\varepsilon_{ijt}   & \mbox{if Group A}\\
\widetilde{\beta_{0}} + \widetilde{\beta_1} \times time_{t}+\varepsilon_{ijt}  & \mbox{if Group B}\\
\end{cases}
(\#eq:ANOVA-lines)
\end{equation}

Presenting the model in this manner makes clear that when treating
different groups, an rm-ANOVA model is able to accommodate non-parallel
lines in each case (different intercepts and slopes per group). In other
words, the rm-ANOVA model ``expects'' a linear relationship between the
covariates and the response, this means that either presented as
Equation @ref(eq:linear-model), Equation @ref(eq:ANOVA-by-group) or
Equation @ref(eq:ANOVA-lines), an rm-ANOVA model is only able to
accommodate linear patterns in the data. If the data show non-linear
behavior, the rm-ANOVA model will approximate this behavior with
non-parallel lines.

\hypertarget{the-linear-mixed-model-case-lmem}{%
\subsubsection{The Linear Mixed Model Case
(LMEM)}\label{the-linear-mixed-model-case-lmem}}

A LMEM is a class of statistical models that incorporates \emph{fixed
effects} to model the relationship between the covariates and the
response, and \emph{random effects} to model subject variability that is
not the primary focus of the study but that might be important to
distinguish \citep{pinheiro2006, west2014}. A LMEM with interaction
between time and treatment for a longitudinal study can be written as:

\begin{equation}
y_{ijt} = \beta_0+ \beta_1 \times time_{t} + \beta_2 \times treatment_{j} + \beta_3 \times time_{t}\times treatment_{j}+\mu_{ij} +\varepsilon_{ijt}\\ 
(\#eq:LMEM)
\end{equation}

When Equation @ref(eq:linear-model) and Equation @ref(eq:LMEM) are
compared, it is easily noticeable that LMEM and rm-ANOVA have the same
construction regarding the \emph{fixed effects} of time and treatment,
but that the LMEM incorporates an additional source of variation (the
term \(\mu_{ij}\)). This term \(\mu_{ij}\) is the one that corresponds
to the \emph{random effect}, accounting for variability in each subject
(subject\(_i\)) within each group (group\(_j\)). The \emph{random}
component can also be understood as used to model some ``noise'' in the
response, but that is intended to be analyzed and disentangled from the
``global noise'' term \(\varepsilon_{ijt}\) from Equation
@ref(eq:linear-model).

For example, if the blood concentration of the drug is measured in
certain subjects in the early hours of the morning while other subjects
are measured in the afternoon, it is possible that the difference in the
collection time introduces some ``noise'' in the data. As the name
suggests, this ``random'' variability needs to be modeled as a variable
rather than as a constant value. The \emph{random effect} \(\mu_{ij}\)
in Equation @ref(eq:LMEM) is assumed to be
\(\mu_{ij} \sim N(0,\sigma^2_\mu)\). In essence, the \emph{random
effect} in a LMEM enables to fit models with different slopes at the
subject-level\citep{pinheiro2006}. However, the expected linear
relationship of the covariates and the response in Equation
@ref(eq:linear-model) and in Equation @ref(eq:LMEM) is essentially the
same, representing a major limitation of LMEMs to fit a non-linear
response.

\hypertarget{covariance-in-rm-anova-and-lmems}{%
\subsection{Covariance in rm-ANOVA and
LMEMs}\label{covariance-in-rm-anova-and-lmems}}

In a longitudinal study there is an expected \emph{covariance} between
repeated measurements on the same subject, and because repeated measures
occur in the subjects within each group, there is a \emph{covariance}
between measurements at each time point within each group. The
\emph{covariance matrix} (also known as the variance-covariance matrix)
is a matrix that captures the variation between and within subjects in a
longitudinal study\citep{wolfinger1996} (For an in-depth analysis of the
covariance matrix see \citep{west2014, weiss2005}).

In the case of an rm-ANOVA analysis, it is typically assumed that the
covariance matrix has a specific construction known as \emph{compound
symmetry} (also known as ``sphericity'' or ``circularity''). Under this
assumption, the between-subject variance and within-subject correlation
are constant across time \citep{weiss2005, geisser1958, huynh1976}.
However, it has been shown that this condition is frequently not
justified because the correlation between measurements tends to change
over time \citep{maxwell2017}; and it is higher between consecutive
measurements \citep{gueorguieva2004, ugrinowitsch2004}. Although
corrections can be made (such as Huyhn-Feldt or
Greenhouse-Geisser)\citep{huynh1976, greenhouse1959} the effectiveness
of each correction is limited because it depends on the size of the
sample,the number of repeated measurements\citep{haverkamp2017}, and
they are not robust if the group sizes are unbalanced
\citep{keselman2001}. Because biomedical longitudinal studies are often
limited in sample size and can have an imbalanced design, the
corrections required to use an rm-ANOVA model may not be able to provide
a reasonable adjustment that makes the model valid.

In the case of LMEMs, one key advantage over rm-ANOVA is that they allow
different structures for the variance-covariance matrix including
exponential, autoregressive of order 1, rational quadratic and others
\citep{pinheiro2006}. Nevertheless, the analysis required to determine
an appropriate variance-covariance structure for the data can be a
challenging process by itself. Overall, the spherical assumption for
rm-ANOVA may not capture the natural variations of the correlation in
the data, and can bias the inferences from the analysis.

\hypertarget{missing-observations}{%
\subsection{Missing observations}\label{missing-observations}}

Missing observations are an issue that arises frequently in longitudinal
studies. In biomedical research, this situation can be caused by reasons
beyond the control of the investigator \citep{molenberghs2004}. Dropout
from patients and attrition or injury in animals are among the reasons
for missing observations. Statistically, missing information can be
classified as \emph{missing at random} (MAR), \emph{missing completely
at random} (MCAR), and \emph{missing not at random} (MNAR)
\citep{weiss2005}. In a MAR scenario, the pattern of the missing
information is related to some variable in the data, but it is not
related to the variable of interest \citep{scheffer2002}. If the data
are MCAR, this means that the missingness is completely unrelated to the
collected information \citep{potthoff2006}, and in the case of MNAR the
missing values are dependent on their value.

An rm-ANOVA model assumes complete observations for all subjects, and
therefore subjects with one or more missing observations are excluded
from the analysis. This is inconvenient because the remaining subjects
might not accurately represent the population, and statistical power is
affected by this reduction in sample size \citep{ma2012}. In the case of
LMEMs, inferences from the model are valid when missing observations in
the data exist that are MAR or MCAR \citep{west2014}. For example, if
attrition occurs in all mice that had lower weights at the beginning of
a chemotherapy response study, the missing data can be considered MAR
because the missigness is unrelated to other variables of interest.

\hypertarget{simulation}{%
\subsection{What do an rm-ANOVA and LMEM fit look like? A visual
representation using simulated data}\label{simulation}}

To visually demonstrate the limitations of rm-ANOVA an LMEMs for
non-linear longitudinal data, this section presents a simulation
experiment of a normally distributed response of two groups of 10
subjects each. An rm-ANOVA model (Equation @ref(eq:linear-model)), and a
LMEM (Equation @ref(eq:LMEM)) are fitted to each group, using
\(\textsf{R}\) \citep{r} and the package \emph{nlme} \citep{nlme}.

Briefly, two cases for the mean responses for each group are considered:
in the first case, the mean response in each group is a linear function
over time with different intercepts and slopes; a negative slope is used
for Group 1 and a positive slope is used for Group 2 (Figure
@ref(fig:l-q-response)A). In the second case, a second-degree polynomial
(quadratic) function is used for the mean response per group: the
quadratic function is concave down for Group 1 and it is concave up for
Group 2 (Figure @ref(fig:l-q-response)C). In both the linear and
quadratic simulated data, the groups start with the same mean value at
the first time point. This is intentional in order to simulate the
expected temporal evolution of some physiological quantity, which is
typical in biomedical experiments where a strong non-linear trend is
present.

Specifically, the rationale for the chosen linear and quadratic
functions is the expectation that a measured response in two treatment
groups is similar in the initial phase of the study, but as therapy
progresses a divergence in the trend of the response indicates a
treatment effect. In other words, Group 1 can be thought as a
``Control'' group and Group 2 as a ``Treatment'' group. From the mean
response per group (linear or quadratic), the variability or ``error''
of individual responses within each group is simulated using a
covariance matrix with compound symmetry (constant variance across
time). Thus, the response per subject in both the linear and quadratic
simulation corresponds to the mean response per group plus the error
(Figure @ref(fig:l-q-response) B,D).

A more comprehensive exploration of the fit of rm-ANOVA and LMEMs for
linear and non-linear longitudinal appears in Figure
@ref(fig:linear-cases-Appendix) and Figure
@ref(fig:quadratic-cases-Appendix) in the Appendix, where simulation
with compound symmetry and independent errors (errors generated from a
normal distribution that are not constant over time) and the plot of
simulated errors, and fitted parameters in presented. We are aware that
the simulated data used in this section present an extreme case that
might not occur frequently in biomedical research, but they are used as
a representation of the consequences of modeling non-linear data with a
linear model such as rm-ANOVA or LMEMs. Of notice, in Section
@ref(longitudinal-GAMs) we use simulated data that does follow reported
trends in the biomedical literature to implement GAMs.

(ref:l-q-response-caption) Simulated responses from two groups with
correlated errors using a LMEM and a rm-ANOVA model. Top row: linear
response, bottom row: quadratic response. A: Simulated linear data with
known mean response (thin lines) and individual responses (points)
showing the dispersion of the data. D: Simulated quadratic data with
known mean response (thin lines) and individual responses (points)
showing the dispersion of the data. B,E: Estimates from the rm-ANOVA
model for the mean group response (linear of quadratic). Points
represent the original raw data. The rm-ANOVA model not only fails to
pick the trend of the quadratic data (D) but also assigns a global
estimate that does not take between-subject variation. C, F: Estimates
from the LMEM in the linear and quadratic case. The LMEM incorporates a
random effect for each subject, but this model and the rm-ANOVA model
are unable to follow the trend of the data and grossly bias the initial
estimates for each group in the quadratic case (bottom row).

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/l-q-response-1}

}

\caption{(ref:l-q-response-caption)}\label{fig:l-q-response}
\end{figure}

The simulation shows that the fits produced by the LMEM and the rm-ANOVA
model are good for linear data, as the predictions for the mean response
are reasonably close to the ``truth'' of the simulated data (Figure
@ref(fig:l-q-response)A). When the linearity and compound symmetry
assumptions are met, the rm-ANOVA model approximates well the global
trend by group (Figure @ref(fig:l-q-response)B). Note that because the
LMEM incorporates \emph{random effects}, is able to provide estimates
for each subject and a ``global'' estimate (Figure
@ref(fig:l-q-response)C).

However, consider the case when the data follows a non-linear trend,
such as the simulated data in Figure @ref(fig:l-q-response)D. Here, the
mean response per group was simulated using a quadratic function, and
errors and individual responses were produced as in Figure
@ref(fig:l-q-response)A. The mean response in the simulated data with
quadratic behavior changes in each group through the timeline, and the
mean value is the same as the initial value by the fifth time point for
each group. Fitting an rm-ANOVA model (Equation @ref(eq:linear-model))
or a LMEM (Equation @ref(eq:LMEM)) to this data produces the fit that
appears in Figure @ref(fig:l-q-response)E, F.

Comparing the fitted responses of the LMEM and the rm-ANOVA models used
in the simulated quadratic data (Figure @ref(fig:l-q-response)E, F)
indicates that the models are not capturing the changes within each
group. Specifically, note that the fitted mean response of both models
shows that the change (increase for Treatment 1 or decrease for
Treatment 2) in the response through time points 2 and 4 is not being
captured. The LMEM is only able to account for between-subject variation
by providing estimates for each subject (Figure
@ref(fig:l-q-response)F), but both models are unable to capture the fact
that the initial values are the same in each group, and instead fit
non-parallel lines that have initial values that are markedly different
from the ``true'' initial values in each case (compare Figure
@ref(fig:l-q-response)D with Figure @ref(fig:l-q-response)E, F). If such
a change has important physiological implications, both rm-ANOVA and
LMEMs omit it from the fitted mean response. Thus, even though the model
correctly detects a divergence between treatment groups, the exact
nature of this difference is not correctly identified, limiting valuable
inferences from the data.

This section has used simulation to better convey the limitations of
linearity and correlation in the response in non-linear data. The models
fitted to the simulated data were an rm-ANOVA model and a LMEM, where
the main issue is the expected linear trend in the response. In the
following section, we present generalized additive models (GAMs) as a
data-driven alternative method to analyze longitudinal non-linear data
that overcomes the linearity assumption.

\hypertarget{GAM-theory}{%
\section{GAMs as a special case of Generalized Linear
Models}\label{GAM-theory}}

\hypertarget{gams-and-basis-functions}{%
\subsection{GAMs and Basis Functions}\label{gams-and-basis-functions}}

Generalized linear models (GLMs) are a family of models (which include
rm-ANOVA and LMEMs) that fit a linear response function to data that may
not have normally distributed errors \citep{nelder1972}. In contrast,
GAMs are a family of regression-based methods for estimating smoothly
varying trends and are a broader class of models that contain the GLM
family as a special case\citep{simpson2018, wood2017, hastie1987}. A GAM
model can be written as:

\begin{equation}
  y_{ijt}=\beta_0+f(x_t\mid \beta_j)+\varepsilon_{ijt}
  (\#eq:GAM)
\end{equation}

Where \(y_{ijt}\) is the response at time \(t\) of subject \(i\) in
group \(j\), \(\beta_0\) is the expected value at time 0, the change of
\(y_{ijt}\) over time is represented by the \emph{smooth function}
\(f(x_t\mid \beta_j)\) with inputs as the covariates \(x_t\) and
parameters \(\beta_j\), and \(\varepsilon_{ijt}\) represents the
residual error.

In contrast to the linear functions used to model the relationship
between the covariates and the response in rm-ANOVA or LMEM, GAMs use
more flexible \emph{smooth functions}. This approach is advantageous as
it does not restrict the model to a linear relationship, although a GAM
can estimate a linear relationship if the data is consistent with a
linear response. One possible set of functions for
\(f(x_t\mid \beta_j)\) that allow for non-linear responses are
polynomials, but a major limitation is that polynomials create a
``global'' fit as they assume that the same relationship exists
everywhere, which can cause problems with inference \citep{beck1998}. In
particular, polynomial fits are known to show boundary effects because
as \(t\) goes to \(\pm \infty\), \(f(x_t \mid \beta_j)\) goes to
\(\pm \infty\) which is almost always unrealistic and causes bias at the
endpoints of the time period.

The smooth functional relationship between the covariates and the
response in GAMs is specified using a semi-parametric relationship that
can be fit within the GLM framework, by using \emph{basis function}
expansions of the covariates and by estimating random coefficients
associated with these basis functions. A \emph{basis} is a set of
functions that spans the mathematical space where the smooths that
approximate \(f(x_t\mid \beta_j)\) exist \citep{simpson2018}. For the
linear model in Equation @ref(eq:linear-model), the basis coefficients
are \(\beta_1\), \(\beta_2\) and \(\beta_3\) and the basis vectors are
\(time_t\), \(treatment_j\) and \(time_t \times treatment_j\). The basis
function then, is the combination of basis coefficients and basis
vectors that map the possible relationship between the covariates and
the response \citep{hefley2017}, which in the case of Equation
@ref(eq:linear-model) is restricted to a linear family of functions. In
the case of Equation @ref(eq:GAM), the basis functions are contained in
the expression \(f(x_t\mid \beta_j)\), which means that the model allows
for non-linear relationships among the covariates.

Splines (cubic, thin plate, etc.) are commonly used \emph{basis
functions}; a cubic spline is a smooth curve constructed from cubic
polynomials joined together in a manner that enforces smoothness, and
thin plate regression splines are an optimized version that work well
with noisy data \citep{wood2017, simpson2018}. Splines have a long
history in solving semi-parametric statistical problems and are often a
default choice to fit GAMs as they are a simple, flexible and powerful
option to obtain smoothness \citep{wegman1983}. Therefore, this
data-driven flexibility in GAMs overcomes the limitation that occurs in
LMEMs and rm-ANOVA when the data is non linear.

To further clarify the concept of basis functions and smooth functions,
consider the simulated response for Group 1 in Figure
@ref(fig:l-q-response)C. The simplest GAM model that can be used to
estimate such response is that of a single smooth term for the time
effect; i.e., a model that fits a smooth to the trend of the group
through time. The timeline can be divided in equally spaced
\emph{knots}, each knot being a region where a different set of basis
functions will be used. Because there are six timepoints for this group,
five knots can be used. The model with five knots to construct the
smooth term means that it will have four basis functions (plus one that
corresponds to the intercept). The choice of basis functions is set
using default values in the package \emph{mgcv} depending on the number
of knots. In Figure @ref(fig:basis-plot)A, the four basis functions (and
the intercept) are shown. Each of the basis functions is composed of six
different points (because there are six points on the timeline). To
control the ``wiggliness'' of the fit, each of the basis functions of
Figure @ref(fig:basis-plot)A is weighted by multiplying it by a
coefficient according to the matrix of Figure @ref(fig:basis-plot)B. The
parameter estimates are penalized (shrunk towards 0) where the penalty
reduces the ``wiggliness'' of the smooth fit to prevent overfitting. A
weak penalty estimate will result in wiggly functions whereas a strong
penalty estimate provides evidence that a linear response is
appropriate.

To get the weighted basis functions, each basis (from Figure Figure
@ref(fig:basis-plot)A) is multiplied by the corresponding coefficients
in Figure @ref(fig:basis-plot)B, thereby increasing or decreasing the
original basis functions. Figure @ref(fig:basis-plot)C shows the
resulting weighted basis functions. Note that the magnitude of the
weighting for the first basis function has resulted in a decrease of its
overall value (because the coefficient for that basis function is less
than 1). On the other hand, the third basis function has roughly doubled
its value. Finally, the weighted basis functions are added at each
timepoint to produce the smooth term. The resulting smooth term for the
effect of \emph{time} is shown in Figure @ref(fig:basis-plot)D (orange
line), along the simulated values per group, which appear as points.

(ref:basis-plot-caption) Basis functions for a single smoother for time
with five knots. A: Basis functions for a single smoother for time for
the simulated data of Group 1 from Figure 2. B: Matrix of basis function
weights. Each basis function is multiplied by a coefficient which can be
positive or negative. The coefficient determines the overall effect of
each basis in the final smoother. C: Weighted basis functions. Each of
the four basis functions of panel A has been weighted by the
corresponding coefficient shown in Panel B. Note the corresponding
increase (or decrease) in magnitude of each weighted basis function. D:
Smoother for time and original data points. The smoother (line) is the
result of the sum of each weighted basis function at each time point,
with simulated values for the group shown as points.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/basis-plot-1}

}

\caption{(ref:basis-plot-caption)}\label{fig:basis-plot}
\end{figure}

\hypertarget{a-bayesian-interpretation-of-gams}{%
\section{A Bayesian interpretation of
GAMs}\label{a-bayesian-interpretation-of-gams}}

Bayes' theorem states that the probability of an event can be calculated
using prior knowledge or belief \citep{mcelreath2018}. In the case of
non-linear data, the belief that the \emph{true} trend of the data is
likely to be smooth rather than ``wiggly'' introduces the concept of a
prior distribution for wiggliness (and therefore a Bayesian view) of
GAMs \citep{wood2017}. GAMs are considered ``empirical'' Bayesian models
because the smoothing parameters are estimated from the data (and not
from a prior distribution as in the ``Full Bayes'' case)
\citep{miller2019}. Moreover, the use of the restricted maximum
likelihood (REML) to estimate the smoothing parameters gives an
empirical estimate of the smooth model \citep{laird1982, pedersen2019}.
Therefore, the confidence intervals calculated for the smooth terms
using the package \emph{mgcv} are considered empirical Bayesian
posterior credible intervals \citep{pedersen2019}, which have good
``frequentist'' coverage (pointwise coverage or ``single point''
coverage), and \emph{across the function} coverage \citep{wood2017}.
This last part means that contrary to a pointwise coverage (where the
coverage of the interval is correct for a single point) the estimated
confidence intervals for the smooths will contain \emph{on average} the
true function of the data 95\% of the time across the entire timeline
(in the case of longitudinal data for which smooths are calculated),
which allows to obtain better inference from the model. In-depth theory
of the Bayesian interpretation of GAMs is beyond the scope of this
paper, but can be found in \citep{miller2019, wood2017, simpson2018} and
\citep{marra2012}. With this brief introduction to the Bayesian
interpretation of GAMs, we henceforth refer to the confidence intervals
for the smooths in GAMs as ``empirical Bayesian'' through the rest of
this paper.

\hypertarget{longitudinal-GAMs}{%
\section{The analyisis of longitudinal biomedical data using
GAMs}\label{longitudinal-GAMs}}

The previous sections provided the basic framework to understand the GAM
framework and how these models are more advantageous to analyze
non-linear longitudinal data when compared to rm-ANOVA or LMEMs. This
section will use simulation to present the practical implementation of
GAMs for longitudinal biomedical data using \(\textsf{R}\) and the
package \texttt{mgcv}. The code for the simulated data and figures, and
a brief guide for model selection and diagnostics appear in the
Appendix.

\hypertarget{simulated-data}{%
\subsection{Simulated data}\label{simulated-data}}

The simulated data is based on the reported longitudinal changes in
oxygen saturation (\(\mbox{StO}_2\)) in subcutaneous tumors that appear
in Figure 3C in \citep{vishwanath2009}. In the paper, diffuse
reflectance spectroscopy was used to quantify \(\mbox{StO}_2\) changes
in both groups at the same time points (days 0, 2, 5, 7 and 10). In the
``Treatment'' group (chemotherapy) an increase in \(\mbox{StO}_2\) is
observed through time, while a decrease is seen in the ``Control''
(saline) group. Following the reported trend, we simulated 10 normally
distributed observations at each time point with a standard deviation
(SD) of 10\% (matching the SD in the original paper). The simulated and
real data appear in Figure @ref(fig:sim-smooth-plot)A and the inset,
respectively.

\hypertarget{an-interaction-gam-for-longitudinal-data}{%
\subsection{An interaction GAM for longitudinal
data}\label{an-interaction-gam-for-longitudinal-data}}

An interaction effect is typically the main interest in longitudinal
biomedical data, as it takes into account treatment, time, and their
combination. In a practical sense, when a GAM is implemented for
longitudinal data, a smooth can be added to the model for the
\emph{time} effect to account for the repeated measures over time.
Although specific methods of how GAMs model correlation structures is a
topic beyond the scope of this paper, it suffices to say that GAMs are
flexible and can handle correlation structures beyond compound symmetry.
A detailed description on basis functions and correlations can be found
in \citep{hefley2017}.

For the data in Figure @ref(fig:sim-smooth-plot), A the main effect of
interest is how \(\mbox{StO}_2\) changes over time for each treatment.
To estimate this, the model incorporates independent smooths for
\emph{Group} and \emph{Day}, respectively. The main thing to consider is
that model syntax accounts for the fact that one of the variables is
numeric (\emph{Day}) and the other is a factor (\emph{Group}). Because
the smooths are centered at 0, the factor variable needs to be specified
as a parametric term in order to identify any differences between the
groups. Using \(\textsf{R}\) and the package \texttt{mgcv} the model
syntax is:

\texttt{m1\ \textless{}-\ gam(StO2\_sim\ \textasciitilde{}\ Group\ +\ s(Day,\ by=Group,\ k=5),\ method=\textquotesingle{}REML,\ data\ =\ dat\_sim)}

This syntax specifies that \texttt{m1} will store the model, and that
the change in the simulated oxygen saturation (\texttt{StO2\_sim}) is
modeled using independent smooths over \texttt{Day} for each
\texttt{Group} (the parenthesis preceded by \texttt{s}) using 5 knots.
The smooth is constructed by default using thin plate regression
splines, but other splines can be used if desired, including Gaussian
process smooths \citep{simpson2018}. The parametric term \texttt{Group}
is added to quantify overall mean differences in the effect of treatment
between groups, and the \texttt{method} chosen to estimate the smoothing
parameters is the restricted maximum likelihood (REML) \citep{wood2017}.
When the smooths are plotted over the raw data, it is clear that the
model has been able to capture the trend of the change of
\(\mbox{StO}_2\) for each group across time (Figure
@ref(fig:sim-smooth-plot)B). Model diagnostics can be obtained using the
\texttt{gam.check} function, and the function \texttt{appraise} from the
package \emph{gratia} \citep{gratia}. A guide for model selection and
diagnostics is in the \protect\hyperlink{workflow}{Appendix}, and an
in-depth analysis can be found in \citep{wood2017} and
\citep{harezlak2018}.

One question that might arise at this point is ``what is the fit that an
rm-ANOVA model produces for the simulated data?''. The rm-ANOVA model,
which corresponds to Equation @ref(eq:linear-model) is presented in
Figure @ref(fig:sim-smooth-plot)C. This is a typical case of model
misspecification: The slopes of each group are different, which would
lead to a \emph{p-value} indicating significance for the treatment and
time effects, but the model is not capturing the changes that occur at
days 2 and between days 5 and 7, whereas the GAM model is able to
reliably estimate the trend over all timepoints (Figure
@ref(fig:sim-smooth-plot)B) .

Because GAMs do not require equally-spaced or complete observations for
all subjects, they are advantageous to analyze longitudinal data where
missingness exists. The rationale behind this is that GAMs are able to
pick the trend in the data even when some observations are missing.
However, this usually causes the resulting smooths to have wider
confidence intervals and less ability to pick certain trends. Consider
the simulated \(\mbox{StO}_2\) values from Figure
@ref(fig:sim-smooth-plot)B. If 40\% of the total observations are
randomly deleted and the same interaction GAM fitted for the complete
dataset is used, the resulting smooths are still able to show a
different trend for each group, but because the empirical Bayesian
credible intervals for the smooths overlap during the first 3 days with
fewer data points, the trend is less pronounced than in the full dataset
(Figure @ref(fig:sim-smooth-plot)D). Although the confidence intervals
have increased for both smooths, the model still shows different trends
with as little as 4 observations per group at certain time points.

(ref:sim-plot-caption) Simulated data and smooths for oxygen saturation
in tumors. A: Simulated data that follows previously reported trends
(inset) in tumors under chemotherapy (Treatment) or saline (Control)
treatment. Simulated data is from a normal distribution with standard
deviation of 10\% with 10 observations per time point. Lines indicate
mean oxygen saturation B: Smooths from the GAM model for the full
simulated data with interaction of Group and Treatment. Lines represent
trends for each group, shaded regions are 95\% confidence intervals. C:
The rm-ANOVA model for the simulated data, which does not capture the
changes in each group over time. D: Smooths for the GAM model for the
simulated data with 40\% of its observations missing. Lines represent
trends for each group, shaded regions are 95\% empirical Bayesian
confidence intervals.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/sim-smooth-plot-1}

}

\caption{(ref:sim-plot-caption)}\label{fig:sim-smooth-plot}
\end{figure}

(ref:pairwise-comp) Pairwise comparisons for smooth terms. A: Pairwise
comparisons for the full dataset. B: Pairwise comparisons for the
dataset with missing observations. Significant differences exist where
the 95\% empirical Bayesian credible interval does not cover 0. In both
cases the effect of treatment is significant after day 3.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/plot-pairwise-comp-1}

}

\caption{(ref:pairwise-comp)}\label{fig:plot-pairwise-comp}
\end{figure}

\hypertarget{GAM-significance}{%
\subsection{Determination of significance in GAMs for longitudinal
data}\label{GAM-significance}}

At the core of a biomedical longitudinal study lies the question of a
significant difference between the effect of two or more treatments in
different groups. Whereas in rm-ANOVA a \emph{post-hoc} analysis is
required to answer such question by calculating some \emph{p-values}
after multiple comparisons, GAMs can use a different approach to
estimate significance. In essence, the idea behind the estimation of
significance in GAMs across different treatment groups is that if the
\emph{difference} between the empirical Bayesian confidence intervals of
the fitted smooths for such groups is non-zero, then a significant
difference exists at that time point(s). The absence of a \emph{p-value}
in this case might seem odd, but the empirical Bayesian confidence
interval interpretation can be conceptualized in the following manner:
Different trends in each group are an indication of an effect by the
treatment. This is what happens for the simulated data in Figure
@ref(fig:sim-smooth-plot)A, where the chemotherapy causes
\(\mbox{StO}_2\) to increase over time.

With this expectation of different trends in each group, computing the
difference between the trends will identify if the observed change is
significant. The difference between groups with similar trends is likely
to yield zero, which would indicate that the treatment is not causing a
change in the response in one of the groups (assuming the other group is
a Control or Reference group).

Consider the calculation of pairwise differences for the smooths in
Figure @ref(fig:sim-smooth-plot)B and Figure @ref(fig:sim-smooth-plot)D.
Figure @ref(fig:plot-pairwise-comp) shows the comparison between each
treatment group for the full and missing datasets. Here, the ``Control''
group is used as the reference to which ``Treatment'' group is being
compared. Of notice, the pairwise comparison has been set on the
response scale (see Appendix for code details), because otherwise the
comparison appears shifted and is not intuitively easy to relate to the
original data.

With this correction in mind, the shaded regions over the confidence
interval (where it does not cover 0) indicate the time interval where
each group has a higher effect than the other. Notice that the shaded
region between days 0 and \(\approx\) 2 for the full dataset indicates
that through that time, the ``Control'' group has higher mean
\(\mbox{StO}_2\), but as therapy progresses the effect is reversed and
by \(\approx\) 3 day it is the ``Treatment'' group the one that on
average, has greater \(\mbox{StO}_2\). This would suggest that the
effect of chemotherapy in the ``Treatment'' group becomes significant
after day 3 for the given model. Moreover, notice that although there is
no actual measurement at day 3, the model is capable of providing an
estimate of when the shift in mean \(\mbox{StO}_2\) occurs.

On the data with missing observations (Figure
@ref(fig:sim-smooth-plot)D), the empirical Bayesian credible intervals
of the smooths overlap between days 0 and 3. Consequently, the smooth
pairwise comparison (Figure @ref(fig:plot-pairwise-comp)B) shows that
there is no evidence of a significant difference between the groups
during that period, but is still able to pick the change on day 3 as the
full dataset smooth pairwise comparison.

In a sense, the pairwise smooth comparison is more informative than a
\emph{post-hoc} \emph{p-value}. For biomedical studies, the smooth
comparison is able to provide an estimate of \emph{when} and by
\emph{how much} a biological process becomes significant. This is
advantageous because it can help researchers gain insight on metabolic
changes and other biological processes that can be worth examining, and
can help refine the experimental design of future studies in order to
obtain measurements at time points where a significant change might be
expected.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Biomedical longitudinal non-linear data is particularly challenging to
analyze due to the likelihood of missing observations and different
correlation structures in the data, which limit the use of rm-ANOVA.
Although LMEMs have started to replace rm-ANOVA as the choice to analyze
biomedical data, both methods yield biased estimates when they are used
to fit non-linear data as we have visually demonstrated in Section
@ref(simulation). This ``model misspecification'' error, also is known
as a ``Type III'' error \citep{dennis2019} is particularly important
because although the \emph{p-value} is the common measure of statistical
significance, the validity of its interpretation is determined by the
agreement of the data and the model. Guidelines for statistical
reporting in biomedical journals exist (the SAMPL guidelines)
\citep{lang2015} but they have not been widely adopted and in the case
of longitudinal data, we consider that researchers would benefit from
reporting a visual assessment of the correspondence between the model
fit and the data, instead of merely relying on a \(R^2\) value.

In this paper we have presented GAMs as a suitable method to analyze
non-linear longitudinal data. It is interesting to note that although
GAMs are a well established method to analyze temporal data in different
fields (among which are palaeoecology, geochemistry, and ecology)
\citep{hefley2017, pedersen2019} they are not routinely used in
biomedical research despite an early publication from Hastie and
Tibshirani that demonstrated their use in medical research
\citep{hastie1995}. This is possibly due to the fact that the theory
behind GAMs can seem very different from that of rm-ANOVA and LMEMs, but
the purpose of Section @ref(GAM-theory) is to demonstrate that at its
core the theory quite simple: Instead of using a linear relationship to
model the response (as rm-ANOVA and LMEMs do), GAMs use basis functions
to build smooths that are capable of following non-linear trends in the
data.

However, from a practical standpoint is equally important to demonstrate
how GAMs are computationally implemented. We have provided an example on
how GAMs can be fitted using simulated data that follows trends reported
in biomedical literature \citep{vishwanath2009} using \(\textsf{R}\) and
the package \emph{mgcv}\citep{wood2017} in Section
@ref(longitudinal-GAMs), while a basic workflow for model selection is
in the \protect\hyperlink{workflow}{Appendix}. One of the features of
GAMs is that their Bayesian interpretation allows to indicate
differences between groups without the need of a \emph{p-value}, and in
turn provide a time-based estimate of shifts in the response that can be
directly tied to biological values as the pairwise smooth comparisons in
Figure @ref(fig:plot-pairwise-comp) indicate. The model is therefore
able to provide an estimate of significant change between the groups at
time points were data was not directly measured even with missing data
exists ( \(\approx\) day 3 in Figure @ref(fig:plot-pairwise-comp) A, B
), which can be used by researchers as feedback on experiment design and
to further evaluate important biological changes in future studies.

We have used \(\textsf{R}\) as the software of choice for this paper
because not only provides a fully developed environment to fit GAMs, but
also eases simulation (which is becoming increasingly used for
exploratory statistical analysis and power calculations) and provides
powerful and convenient methods of visualization, which are key aspects
that biomedical researchers might need to consider to make their work
reproducible. In this regard, reproducibility is still an issue in
biomedical research \citep{begley2015, weissgerber2018}, but it is
becoming apparent that what other disciplines have experienced in this
aspect is likely to impact sooner rather than later this field.
Researchers need to plan on how they will make their data, code, and any
other materials open and accessible as more journals and funding
agencies recognize the importance and benefits of open science in
biomedical research. We have made all the data and code used in this
paper accessible, and we hope that this will encourage other researchers
to do the same with future projects.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

We have presented GAMs as a method to analyze longitudinal biomedical
data. Future directions of this work will include simulation-based
estimations of statistical power using GAMs, as well as demonstrating
the prediction capabilities of these models using large datasets. By
making the data and code used in this paper accessible, we hope to
address the need of creating and sharing reproducible work in biomedical
research.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

This work was supported by the National Science Foundation Career Award
(CBET 1751554, TJM) and the Arkansas Biosciences Institute.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{bibliography}{%
\section{Bibliography}\label{bibliography}}

\hypertarget{appendix}{%
\section{(APPENDIX)}\label{appendix}}

\hypertarget{code-for-manuscript-data}{%
\section{Code for Manuscript data}\label{code-for-manuscript-data}}

This section presents the code used to generate figures, models and
simulated data from Sections 3 and 4 from the main manuscript.

\hypertarget{compound-symmetry-and-independent-errors-in-linear-and-quadratic-responses}{%
\subsection{Compound symmetry and independent errors in linear and
quadratic
responses}\label{compound-symmetry-and-independent-errors-in-linear-and-quadratic-responses}}

This section simulated linear and quadratic data in the same manner as
in Section @ref(simulation). The linear simulations using Figure
@ref(fig:linear-cases-Appendix) show in panels A and D the simulated
mean responses and individual data points. Panels C and G show a visual
interpretation of ``correlation'' in the responses: In panel C, subjects
that have a value of the random error \(\varepsilon\) either above or
below the mean group response are more likely to have other observations
that follow the same trajectory, thereby demonstrating correlation in
the response. In panel G,because the errors are independent, there is no
expectation that responses are likely to follow a similar pattern.
Panels D and H show the predictions from the rm-ANOVA model.

The following code produces a more comprehensive exploration of Figure
@ref(fig:l-q-response) in the main manuscript.

\begin{verbatim}
##########Section for calculations###########

## Example with linear response

#This function simulates data using a linear or quadratic mean response and each with correlated
#or uncorrelated errors. Each group has a different slope/concavity.
example <- function(n_time = 6, #number of time points
                    fun_type = "linear", #type of response
                    error_type = "correlated") {

  if (!(fun_type %in% c("linear", "quadratic")))
    stop('fun_type must be either "linear", or "quadratic"')
  if (!(error_type %in% c("correlated", "independent")))
    stop('fun_type must be either "correlated", or "independent"')

  x <- seq(1,6, length.out = n_time)

  #Create mean response matrix: linear or quadratic
  mu <- matrix(0, length(x), 2)
  # linear response
  if (fun_type == "linear") {
    mu[, 1] <- - (0.25*x)+2  
    mu[, 2] <- 0.25*x+2
  } else {
    # quadratic response (non-linear)

    mu[, 1] <-  -(0.25 * x^2) +1.5*x-1.25
    mu[, 2] <- (0.25 * x^2) -1.5*x+1.25
  }

  #create an array where individual observations per each time point for each group are to be stored. Currently using 10 observations per timepoint
  y <- array(0, dim = c(length(x), 2, 10))

  #Create array to store the "errors" for each group at each timepoint. The "errors" are the 
  #between-group variability in the response.
  errors <- array(0, dim = c(length(x), 2, 10))
  #create an array where 10 observations per each time point for each group are to be stored

  #The following cycles create independent or correlated responses. To each value of mu (mean response per group) a randomly generated error (correlated or uncorrelated) is added and thus the individual response is created.
  if (error_type == "independent") {
    ## independent errors
    for (i in 1:2) {
      for (j in 1:10) {
        errors[, i, j] <- rnorm(6, 0, 0.25)
        y[, i, j] <- mu[, i] + errors[, i, j]
      }
    }
  } else {
    for (i in 1:2) {     # number of treatments
      for (j in 1:10) {  # number of subjects
        # compound symmetry errors: variance covariance matrix
        errors[, i, j] <- rmvn(1, rep(0, length(x)), 0.1 * diag(6) + 0.25 * matrix(1, 6, 6))
        y[, i, j] <- mu[, i] + errors[, i, j]
      }
    }
  }

  ## subject random effects

  ## visualizing the difference between independent errors and compound symmetry
  ## why do we need to account for this -- overly confident inference

#labeling y and errors  
  dimnames(y) <- list(time = x, 
                      treatment = 1:2, 
                      subject = 1:10)

  dimnames(errors) <- list(time = x, 
                           treatment = 1:2, 
                           subject = 1:10)

  #labeling the mean response
  dimnames(mu) <- list(time = x, 
                       treatment = 1:2)

  #convert y, mu and errors to  dataframes with time, treatment and subject columns
  dat <- as.data.frame.table(y, 
                             responseName = "y")
  dat_errors <- as.data.frame.table(errors, 
                                    responseName = "errors")
  dat_mu <- as.data.frame.table(mu, 
                                responseName = "mu")

  #join the dataframes to show mean response and errors per subject
  dat <- left_join(dat, dat_errors, 
                   by = c("time", "treatment", "subject"))
  dat <- left_join(dat, dat_mu, 
                   by = c("time", "treatment"))
  #add time
  dat$time <- as.numeric(as.character(dat$time))
  #label subjects per group
  dat <- dat %>%
    mutate(subject = factor(paste(subject, 
                                  treatment, 
                                  sep = "-")))

  ## repeated measures ANOVA

  fit_anova <- lm(y ~ time + treatment + time * treatment, data = dat)

#LMEM: time and treatment interaction model, compound symmetry 
  fit_lme <- lme(y ~ treatment + time + treatment:time,
                 data = dat,
                 random = ~ 1 | subject,
                 correlation = corCompSymm(form = ~ 1 | subject)
  )

  #create a prediction frame where the model can be used for plotting purposes
  pred_dat <- expand.grid(
    treatment = factor(1:2), 
    time = unique(dat$time)
  )

  #add model predictions to the dataframe that has the simulated data
  dat$pred_anova <- predict(fit_anova)
  dat$pred_lmem <- predict(fit_lme)

  #return everything in a list
  return(list(
    dat = dat,
    pred_dat = pred_dat,
    fit_anova=fit_anova,
    fit_lme = fit_lme
  ))
}
##################Section for plotting#################################
#######################################################################
#This function will create the plots for either a "linear" or "quadratic" response

plot_example <- function(sim_dat) {
  ## Plot the simulated data (scatterplot)

  p1 <- sim_dat$dat %>%
    ggplot(aes(x = time, 
               y = y, 
               group = treatment, 
               color = treatment)
           ) +
    geom_point(show.legend=FALSE) +
    labs(y='response')+
    geom_line(aes(x = time, 
                  y = mu, 
                  color = treatment),
              show.legend=FALSE) +
    theme_classic() +
    theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    thm

  #plot the simulated data with trajectories per each subject
  p2 <- sim_dat$dat %>%
    ggplot(aes(x = time, 
               y = y, 
               group = subject, 
               color = treatment)
           ) +
    geom_line(aes(size = "Subjects"),
              show.legend = FALSE) +
    # facet_wrap(~ treatment) +
    geom_line(aes(x = time, 
                  y = mu, 
                  color = treatment,
                  size = "Simulated Truth"), 
              lty = 1,show.legend = FALSE) +
    labs(y='response')+
    scale_size_manual(name = "Type", values=c("Subjects" = 0.5, "Simulated Truth" = 3)) +
    theme_classic()+
     theme(plot.title = element_text(size = 30, 
                                face = "bold"),
     text=element_text(size=30))+
    thm

  #plot the errors
   p3 <- sim_dat$dat %>%
    ggplot(aes(x = time, 
               y = errors,
               group = subject, 
               color = treatment)) +
    geom_line(show.legend=FALSE) +
     labs(y='errors')+
     theme_classic()+
     theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    thm

   #plot the model predictions for rm-ANOVA
  p4 <- ggplot(sim_dat$dat, 
               aes(x = time, 
                   y = y, 
                   color = treatment)) +
    geom_point(show.legend=FALSE)+
    labs(y='response')+
    geom_line(aes(y = predict(sim_dat$fit_anova), 
                  group = subject, size = "Subjects"),show.legend = FALSE) +
    geom_line(data = sim_dat$pred_dat, 
              aes(y = predict(sim_dat$fit_anova, 
                              level = 0, 
                              newdata = sim_dat$pred_dat), 
                  size = "Population"),
              show.legend=FALSE) +
    guides(color = guide_legend(override.aes = list(size = 2)))+
    scale_size_manual(name = "Predictions", 
                      values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_classic() +
    theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    thm

   #plot the LMEM predictions
  p5 <- ggplot(sim_dat$dat, 
               aes(x = time, 
                   y = y, 
                   color = treatment)) +
    geom_point()+
    labs(y='response')+
    geom_line(aes(y = predict(sim_dat$fit_lme), 
                  group = subject, size = "Subjects")) +
    geom_line(data = sim_dat$pred_dat, 
              aes(y = predict(sim_dat$fit_lme, 
                              level = 0, 
                              newdata = sim_dat$pred_dat), 
                  size = "Population")) +
    guides(color = guide_legend(override.aes = list(size = 2)))+
    scale_size_manual(name = "Predictions", 
                      values=c("Subjects" = 0.5, "Population" = 3)) +
    theme_classic() +
    theme(plot.title = element_text(size = 30, 
                                  face = "bold"),
        text=element_text(size=30))+
    thm

  return((p1+p3+p2+p4+p5)+plot_layout(nrow=1)+plot_annotation(tag_levels = 'A'))

}

txt<-18

#Store each plot in a separate object
A1<-plot_example(example(fun_type = "linear", error_type = "correlated"))

B1<-plot_example(example(fun_type = "linear", error_type = "independent"))

C1<-plot_example(example(fun_type = "quadratic", error_type = "correlated"))

D1<-plot_example(example(fun_type = "quadratic", error_type = "independent")) 
\end{verbatim}

(ref:linear-cases-appendix) Simulated linear responses from two groups
with correlated (top row) or independent (bottom row) errors using a
rm-ANOVA model and a LMEM. A, F:Simulated data with known mean response
and individual responses (points) showing the dispersion of the data.
B,G: Generated errors showing the difference in the behavior of
correlated and independent errors. C,H: Simulated data with thin lines
representing individual trajectories. D,I: Estimations from the rm-ANOVA
model for the mean group response. E, J: Estimations from the LMEM for
the mean group response and individual responses (thin lines). In all
panels, thick lines are the predicted mean response per group, thin
lines are the random effects for each subject and points represent the
original raw data. Both rm-ANOVA and the LMEM are able to capture the
trend of the data.

\begin{figure}
\centering
\includegraphics{Full_document_SAGE_files/figure-latex/linear-cases-Appendix-1.pdf}
\caption{(ref:linear-cases-appendix)}
\end{figure}

For the quadratic response case, Figure
@ref(fig:quadratic-cases-Appendix) shows the simulated responses using
compound symmetry and independent errors.

(ref:quadratic-caption-appendix) Simulated quadratic responses from two
groups with correlated (top row) or independent (bottom row) errors
using a rm-ANOVA model and a LMEM. A, F:Simulated data with known mean
response and individual responses (points) showing the dispersion of the
data. B,G: Generated errors showing the difference in the behavior of
correlated and independent errors. C,H: Simulated data with thin lines
representing individual trajectories. D,I: Estimations from the rm-ANOVA
model for the mean group response. E, J: Estimations from the LMEM for
the mean group response and individual responses (thin lines). In all
panels, thick lines are the predicted mean response per group, thin
lines are the random effects for each subject and points represent the
original raw data. Both rm-ANOVA and the LMEM are not able to capture
the changes in each group over time.

\begin{figure}
\centering
\includegraphics{Full_document_SAGE_files/figure-latex/quadratic-cases-Appendix-1.pdf}
\caption{(ref:quadratic-caption-appendix)}
\end{figure}

\hypertarget{basis-functions-and-gams}{%
\subsection{Basis functions and GAMs}\label{basis-functions-and-gams}}

This code produces Figure @ref(fig:basis-plot) from the main manuscript.
Briefly, a non-linear (quadratic) response is simulated a gam model is
fitted and the basis are extracted in order to explain how the smooth is
constructed. The code for data simulation is used again here for the
sake of keeping the same structure, although the data can be simulated
in a more simple fashion.

\begin{verbatim}
#generate the response: the same initial procedure from the previous section to simulate
#the response
set.seed(1)
n_time = 6
 x <- seq(1,6, length.out = n_time)
 mu <- matrix(0, length(x), 2)
 mu[, 1] <-  -(0.25 * x^2) +1.5*x-1.25 #mean response
 mu[, 2] <- (0.25 * x^2) -1.5*x+1.25 #mean response
 y <- array(0, dim = c(length(x), 2, 10))
 errors <- array(0, dim = c(length(x), 2, 10))
 for (i in 1:2) {     # number of treatments
     for (j in 1:10) {  # number of subjects
         # compound symmetry errors
         errors[, i, j] <- rmvn(1, rep(0, length(x)), 0.1 * diag(6) + 0.25 * matrix(1, 6, 6))
         y[, i, j] <- mu[, i] + errors[, i, j]
     }
 }

 #label each table
  dimnames(y) <- list(time = x, treatment = 1:2, subject = 1:10)
 dimnames(errors) <- list(time = x, treatment = 1:2, subject = 1:10)
 dimnames(mu) <- list(time = x, treatment = 1:2)

 #Convert to dataframes with subject, time and group columns
 dat <- as.data.frame.table(y, responseName = "y")
 dat_errors <- as.data.frame.table(errors, responseName = "errors")
 dat_mu <- as.data.frame.table(mu, responseName = "mu")
 dat <- left_join(dat, dat_errors, by = c("time", "treatment", "subject"))
 dat <- left_join(dat, dat_mu, by = c("time", "treatment"))
 dat$time <- as.numeric(as.character(dat$time))

 #label subject per group
 dat <- dat %>%
     mutate(subject = factor(paste(subject, treatment, sep = "-")))

 #extract  "Group 1" to fit the GAM
  dat<-subset(dat,treatment==1)
 #keep just the response and timepoint columns
   dat<-dat[,c('y','time')]

   #GAM model of time, 5 knots
gm<-gam(y~s(time,k=5),data=dat)

#model_matrix (also known as) 'design matrix'
#will contain the smooths used to create  model 'gm'
model_matrix<-as.data.frame(predict(gm,type='lpmatrix'))

time<-c(1:6)

basis<-model_matrix[1:6,] #extracting basis (because the values are repeated after every 6 rows)
#basis<-model_matrix[1:6,-1] #extracting basis
colnames(basis)[colnames(basis)=="(Intercept)"]<-"s(time).0"
basis<-basis %>% #pivoting to long format
  pivot_longer(
    cols=starts_with("s")
  )%>%
  arrange(name) #ordering

#length of dataframe to be created: number of knots by number of timepoints (minus 1 for the intercept that we won't plot)
ln<-6*(length(coef(gm)))

basis_plot<-data.frame(Basis=integer(ln),
                       value_orig=double(ln),
                       time=integer(ln),
                       cof=double(ln)
)

basis_plot$time<-rep(time) #pasting timepoints
basis_plot$Basis<-factor(rep(c(1:5),each=6)) #pasting basis number values
basis_plot$value_orig<-basis$value #pasting basis values
basis_plot$cof<-rep(coef(gm)[1:5],each=6) #pasting coefficients
basis_plot<-basis_plot%>%
  mutate(mod_val=value_orig*cof) #the create the predicted values the bases need to be 
#multiplied by the coefficients

#creating labeller to change the labels in the basis plots

basis_names<-c(
  `1`="Intercept",
  `2`="1",
  `3`="2",
  `4`="3",
  `5`="4"
)

#calculating the final smooth by aggregating the basis functions

smooth<-basis_plot%>% 
  group_by(time)%>%
  summarize(smooth=sum(mod_val))

#original basis
sz<-1
p11<-ggplot(basis_plot,
            aes(x=time,
                y=value_orig,
                colour=as.factor(Basis)
                )
            )+
  geom_line(size=sz,
            show.legend=FALSE)+
  geom_point(size=sz+1,
             show.legend = FALSE)+
  labs(y='Basis functions')+
  facet_wrap(~Basis,
             labeller = as_labeller(basis_names)
             )+
  theme_classic()+
  thm

#penalized basis
p12<-ggplot(basis_plot,
            aes(x=time,
                y=mod_val,
                colour=as.factor(Basis)
                )
            )+
  geom_line(show.legend = FALSE,
            size=sz)+
  geom_point(show.legend = FALSE,
             size=sz+1)+
  labs(y='Penalized \n basis functions')+
  scale_y_continuous(breaks=seq(-1,1,1))+
  facet_wrap(~Basis,
             labeller=as_labeller(basis_names)
             )+
  theme_classic()+
  thm

#heatmap of the coefficients
x_labels<-c("Intercept","1","2","3","4")
p13<-ggplot(basis_plot,
            aes(x=Basis,
                y=Basis))+
  geom_tile(aes(fill = cof), 
            colour = "black") +
    scale_fill_gradient(low = "white",
                        high = "#B50A2AFF")+ #color picked from KikiMedium
  labs(x='Basis',
       y='Basis')+
  scale_x_discrete(labels=x_labels)+
  geom_text(aes(label=round(cof,2)),
            size=7,
            show.legend = FALSE)+
  theme_classic()+
  theme(legend.title = element_blank())

#plotting simulated datapoints and smooth term
p14<-ggplot(data=dat,
            aes(x=time,y=y))+
  geom_point(size=sz+1)+
  labs(y='Simulated \n response')+
  geom_line(data=smooth,
            aes(x=time,
                y=smooth),
            color="#6C581DFF",
            size=sz+1)+
  theme_classic()

#Combining all
b_plot<-p11+p13+p12+p14+plot_annotation(tag_levels='A')&
  theme(
     text=element_text(size=18)
     )
\end{verbatim}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/basis-plot-appendix-1}

}

\caption{(ref:basis-plot-caption)}\label{fig:basis-plot-appendix}
\end{figure}

\hypertarget{tumor-data-simulation}{%
\section{Longitudinal biomedical data simulation and
GAMs}\label{tumor-data-simulation}}

This section describes how to fit GAMs to longitudinal data using
simulated data. First, data is simulated according to Section
@ref(longitudinal-GAMs), where reported data of oxygen saturation
(\(\mbox{StO}_2\)) in tumors under either chemotherapy or saline control
is used as a starting point to generate individual responses in each
group.

\begin{verbatim}
set.seed(1)
#Dataframe that contains the original reported trends
dat<-tibble(StO2=c(4,27,3,2,0.5,7,4,50,45,56),
            Day=rep(c(0,2,5,7,10),times=2),
            Group=as.factor(rep(c("Control","Treatment"),each=5))
            )

## plot the mean response
f1<-ggplot(dat, 
           aes(x = Day, 
               y = StO2, 
               color = Group)) +
    geom_line(size=1,
              show.legend = FALSE)+
    geom_point(show.legend = FALSE,
               size=1.5,
               alpha=0.5)+
  labs(y=expression(paste(StO[2],
                          ' (real)')))+
  theme_classic()+
  thm+
    scale_x_continuous(breaks=c(0,5,10))+
    scale_y_continuous(breaks=c(0,40))+
  plot_layout(tag_level = 'new')+
  theme(
    plot.background = element_rect(fill = "transparent", 
                                   color = NA),
    axis.text=element_text(size=14)
  )

#This function simulates data for the tumor data using default parameters of 10 observations per time point,and Standard deviation (sd) of 5%.
#Because physiologically StO2 cannot go below 0%, data is  generated with a cutoff value of 0.0001 (the "StO2_sim")

simulate_data <- function(dat, n = 10, sd = 5) {
    dat_sim <- dat %>%
        slice(rep(1:n(), each = n)) %>%
        group_by(Group, Day) %>%
        mutate(
               StO2_sim = pmax(rnorm(n, StO2, sd), 0.0001),
               subject=rep(1:10),
               subject=factor(paste(subject, Group, sep = "-"))
               ) %>%
        ungroup()

    return(dat_sim)
}

#subject = factor(paste(subject, treatment, sep = "-")))
n <- 10 #number of observations
sd <- 10 #approximate sd from paper
df <- 6
dat_sim <- simulate_data(dat, n, sd)

#plotting simulated data
f2<-ggplot(dat_sim, 
           aes(x = Day, 
               y = StO2_sim, 
               color = Group)) +
    geom_point(show.legend=FALSE,
               size=1.5,
               alpha=0.5)+
    stat_summary(aes(y = StO2_sim,
                     group=Group), 
                 fun=mean, geom="line",
                 size=1,
                 show.legend = FALSE)+
  labs(y=expression(atop(StO[2],
                         '(simulated)')))+
  theme_classic()+
  theme(
    axis.text=element_text(size=22)
  )+
  thm+
    scale_x_continuous(breaks=c(0,2,5,7,10))
\end{verbatim}

\hypertarget{workflow}{%
\subsection{A basic Workflow for GAMs}\label{workflow}}

This section shows a basic workflow to fit a series of increasingly
complex GAMs to the simulated data from the previous section. Graphical
and parameter diagnostics for goodness of fit are discussed, as well as
model comparison via AIC (Aikake Information Criterion).

\hypertarget{first-model}{%
\subsubsection{First model}\label{first-model}}

The first model fitted to the data is one that only accounts for
different smooths by day. The model syntax specifies that
\texttt{gam\_00} is the object that will contain all the model
information, and that the model attempts to explain changes in
\texttt{StO2\_sim} (simulated \(\mbox{StO}_2\)) using a smooth per
\texttt{Day}. The model will use 5 knots (\texttt{k=5}) for the smooth.
The smooth is constructed by default using thin plate regression
splines. The smoothing parameter estimation method used is the
restricted maximum likelihood (\texttt{REML}).

\begin{verbatim}
gam_00<-gam(StO2_sim ~ s(Day, k = 5),
            method='REML',
            data  = dat_sim)
\end{verbatim}

To obtain model diagnostics, two methodologies are to be used: 1)
graphical diagnostics, and 2) a model check. In the first case, the
functions \texttt{appraise} and \texttt{draw} from the package
\emph{gratia} can be used to obtain a single output with all the
graphical diagnostics. For model check, the functions \texttt{gam.check}
and \texttt{summary} from \emph{mgcv} provide detailed information about
the model fit and its parameters.

(ref:first-GAM-caption) Graphical diagnostics for the first GAM model.
Left: Graphical diagnostics provided by the function \texttt{appraise}
from the package \emph{gratia}. Right: Fitted smooth for the model,
provided by the function \texttt{draw}.

\hypertarget{graphical-diagnostics}{%
\paragraph{Graphical diagnostics}\label{graphical-diagnostics}}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/first-GAM-diag-1}

}

\caption{(ref:first-GAM-caption)}\label{fig:first-GAM-diag}
\end{figure}

From the output of the function \texttt{appraise} in Figure
@ref(fig:first-GAM-diag), the major indicators of concern about the
model are the QQ plot of residuals and the histogram of residuals. The
QQ plot shows that the errors are not reasonably located along the
45\(^{\circ}\) line (which indicates normality), as there are multiple
points that deviate from the trend, specially in the tails. The
histogram also shows that the variation (residuals) is not following the
assumption of a normal distribution.

The \texttt{draw} function permits to plot the smooths as
\texttt{ggplot2} objects, which eases subsequent manipulation, if
desired. Because model \texttt{gam\_00} specifies only one smooth for
the time covariate (Day), the plot only contains only one smooth. Note
that the smooth shows an almost linear profile.

\hypertarget{model-check}{%
\paragraph{Model check}\label{model-check}}

\begin{verbatim}
#need to add figure number and caption
gam.check(gam_00)
\end{verbatim}

\includegraphics{Full_document_SAGE_files/figure-latex/first-GAM-check-1.pdf}

\begin{verbatim}
## 
## Method: REML   Optimizer: outer newton
## full convergence after 5 iterations.
## Gradient range [-0.0003727881,-6.621452e-07]
## (score 444.0118 & scale 450.6638).
## Hessian positive definite, eigenvalue range [0.3881695,49.00676].
## Model rank =  5 / 5 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##          k'  edf k-index p-value    
## s(Day) 4.00 2.11    0.36  <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

\begin{verbatim}
summary(gam_00)
\end{verbatim}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## StO2_sim ~ s(Day, k = 5)
## 
## Parametric coefficients:
##             Estimate Std. Error t value Pr(>|t|)    
## (Intercept)   22.967      2.123   10.82   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Approximate significance of smooth terms:
##          edf Ref.df     F  p-value    
## s(Day) 2.114  2.565 7.633 0.000517 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## R-sq.(adj) =  0.153   Deviance explained = 17.2%
## -REML = 444.01  Scale est. = 450.66    n = 100
\end{verbatim}

Special attention must be paid to the `k-index' from \texttt{gam.check}.
This parameter indicates if the basis dimension of the smooth is
adequate, i.e., it checks that the basis used to create the smooth are
adequate to capture the trends in the data. If the model is not
adequately capturing the trens in the data, this is indicated by a low
k-index value (\textless1). From the output, it can be seen that the
k-index is 0.36, which indicates that the model is not capturing the
variability in the data. The \texttt{edf} (effective degrees of freedom)
is an indicator of the complexity of the smooth. Here the complexity of
the smooth is comparable to that of a 4th degree polynomial.

From the \texttt{summary} function, information about the assumed
distribution of the errors (Gaussian in this case) and the link function
can be obtained. The link function is `identity' as the model does not
make any transformation on the predictors. The `significance of smooth
terms' \emph{p-value} indicates if each smooth is adding significance to
the model. Here, the \emph{p-value} is low but we have seen that there
are issues with the model from the previous outputs. Finally, the
`deviance explained' indicates how much of the data the model is able to
capture, which in this case corresponds to \(\approx\) 17\%.

\hypertarget{second-model}{%
\subsubsection{Second model}\label{second-model}}

The major flaw of \texttt{gam\_00} is that this model is not taking into
account the fact that the data is nested in groups. The next iteration
is a model where a different smooth of time (Day) is assigned for each
group using \texttt{by=Group} in the model syntax.

\begin{verbatim}
gam_01<-gam(StO2_sim ~ s(Day, by=Group,k = 5),
            method='REML',
            data  = dat_sim)

gam.check(gam_01)
\end{verbatim}

\includegraphics{Full_document_SAGE_files/figure-latex/second-GAM-1.pdf}

\begin{verbatim}
## 
## Method: REML   Optimizer: outer newton
## full convergence after 7 iterations.
## Gradient range [-5.51754e-05,2.671715e-06]
## (score 423.3916 & scale 280.8777).
## Hessian positive definite, eigenvalue range [0.3162258,48.5557].
## Model rank =  9 / 9 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##                         k'  edf k-index p-value    
## s(Day):GroupControl   4.00 3.39    0.43  <2e-16 ***
## s(Day):GroupTreatment 4.00 3.23    0.43  <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
\end{verbatim}

\begin{verbatim}
summary(gam_01)
\end{verbatim}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## StO2_sim ~ s(Day, by = Group, k = 5)
## 
## Parametric coefficients:
##             Estimate Std. Error t value Pr(>|t|)    
## (Intercept)   22.967      1.676    13.7   <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Approximate significance of smooth terms:
##                         edf Ref.df      F p-value    
## s(Day):GroupControl   3.392  3.794  3.817  0.0304 *  
## s(Day):GroupTreatment 3.229  3.682 21.174  <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## R-sq.(adj) =  0.472   Deviance explained = 50.8%
## -REML = 423.39  Scale est. = 280.88    n = 100
\end{verbatim}

Diagnostics for this model indicate that the k-index is still below 1
(0.43 from \texttt{gam.check}), and that the residuals are still not
following a normal distribution (Figure @ref(fig:second-GAM-diag)).
Moreover, the smooths (plotted via the \texttt{draw()} function) appear
with a fairly linear profile, which indicates they are still not
capturing the trends observed in the data. From \texttt{summary()}, the
deviance explained by the model is \(\approx\) 51\%.

(ref:second-GAM-caption) Graphical diagnostics for the second GAM model.
Left: Graphical diagnostics provided by the function \texttt{appraise}
from the package \emph{gratia}. Right: Fitted smooth for the model,
provided by the function \texttt{draw}.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/second-GAM-diag-1}

}

\caption{(ref:second-GAM-caption)}\label{fig:second-GAM-diag}
\end{figure}

\hypertarget{third-model}{%
\subsubsection{Third model}\label{third-model}}

Model \texttt{gam\_00} was built for didactic purposes to cover the
simplest case, but it does not account for the nesting of the data by
Group, which is apparent from the type of smooth fitted, the model
diagnostics, and, the low variance explained by the model. On the other
hand, \texttt{gam\_01} takes into account the nesting within each group
and provides better variance explanation, but as indicated in Section
@ref(longitudinal-GAMs), in order to differentiate between each group a
parametric term needs to be added to the model for the interaction of
\emph{Day} and \emph{Group}.

\begin{verbatim}
#GAM for StO2

m1 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5),
            method='REML',
            data  = dat_sim)

gam.check(m1)
\end{verbatim}

\includegraphics{Full_document_SAGE_files/figure-latex/final-model-Appendix-1.pdf}

\begin{verbatim}
## 
## Method: REML   Optimizer: outer newton
## full convergence after 10 iterations.
## Gradient range [-8.164307e-08,1.500338e-08]
## (score 355.8554 & scale 64.53344).
## Hessian positive definite, eigenvalue range [1.174841,48.08834].
## Model rank =  10 / 10 
## 
## Basis dimension (k) checking results. Low p-value (k-index<1) may
## indicate that k is too low, especially if edf is close to k'.
## 
##                         k'  edf k-index p-value
## s(Day):GroupControl   4.00 3.87    1.02    0.59
## s(Day):GroupTreatment 4.00 3.88    1.02    0.54
\end{verbatim}

\begin{verbatim}
summary(m1)
\end{verbatim}

\begin{verbatim}
## 
## Family: gaussian 
## Link function: identity 
## 
## Formula:
## StO2_sim ~ Group + s(Day, by = Group, k = 5)
## 
## Parametric coefficients:
##                Estimate Std. Error t value Pr(>|t|)    
## (Intercept)       9.084      1.136   7.996 4.09e-12 ***
## GroupTreatment   27.766      1.607  17.282  < 2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Approximate significance of smooth terms:
##                         edf Ref.df     F p-value    
## s(Day):GroupControl   3.873  3.990 17.57  <2e-16 ***
## s(Day):GroupTreatment 3.879  3.991 89.33  <2e-16 ***
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## R-sq.(adj) =  0.879   Deviance explained = 88.9%
## -REML = 355.86  Scale est. = 64.533    n = 100
\end{verbatim}

The resulting model is \texttt{m1}, which is the model fitted in the
main manuscript. By using \texttt{appraise()} and \texttt{draw} on this
model (Figure @ref(fig:final-GAM-diag)) we see that the trend on the QQ
plot has improved, the histogram of the residuals appears to be
reasonably distributed, and the smooths are capturing the trend of the
data within each group. From \texttt{gam.check}, the k-index is now at
an acceptable value (\(\approx\) 1.02), and \texttt{summary} now
indicates that the model is able to capture 89\% of the variance in the
data.

(ref:final-GAM-caption) Graphical diagnostics for the final GAM model.
Left: Graphical diagnostics provided by the function \texttt{appraise}
from the package \emph{gratia}. Right: Fitted smooths for the model,
provided by the function \texttt{draw}.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/final-GAM-diag-1}

}

\caption{(ref:final-GAM-caption)}\label{fig:final-GAM-diag}
\end{figure}

\hypertarget{comparing-models-via-aic}{%
\subsubsection{Comparing models via
AIC}\label{comparing-models-via-aic}}

One final comparison that can be made for model selection involves the
use of the Aikake Information Criterion (AIC). This metric is used to
estimate information loss, which we want to minimize with an appropriate
model. Therefore, when 2 or more models are compared, the model with
lower AIC is preferred. In R, the comparison is done using the
\texttt{AIC} function.

\begin{verbatim}
AIC(gam_00,gam_01,m1)
\end{verbatim}

\begin{verbatim}
##               df      AIC
## gam_00  4.564893 900.8257
## gam_01  9.476137 858.6051
## m1     10.980983 712.2067
\end{verbatim}

The output in this case is expected: model \texttt{m1} has a lower AIC
(712.46) whereas the initial two models have higher AICs (900 and 858).
The AIC should not be considered as the only estimator of model quality,
instead to be used as complimentary information to the graphical
diagnostics and model checks described above.

\hypertarget{pairwise-comparisons-of-smooth-confidence-intervals}{%
\paragraph{Pairwise comparisons of smooth confidence
intervals}\label{pairwise-comparisons-of-smooth-confidence-intervals}}

The estimation of significant differences between each treatment group
can be achieved via pairwise comparisons of the smooth confidence
intervals as described in section @ref(GAM-significance). In this case,
the ``design matrix'' is used to estimate the pairwise comparisons (see
main manuscript for details and associated references). Briefly, the
``design matrix'' (also known as the ``Xp matrix'') from the selected
model (\texttt{m1}) is used to calculate a 95\% confidence interval of
the difference between the smooth terms for each group. This approach
allows to estimate the time intervals where a significant difference
exists between the groups (confidence interval above or below 0).
\textbf{All pairwise comparisons in this paper have been centered at the
response scale to ease interpretation }.

\begin{verbatim}
##Pairwise comparisons
pdat <- expand.grid(Day = seq(0, 10, length = 400),
                    Group = c('Control', 'Treatment'))

##matrix that contains the basis functions evaluated at the points in pdat
    xp <- predict(m1, newdata = pdat, type = 'lpmatrix')

#Find columns in xp where the name contains "Control"
    c1 <- grepl('Control', colnames(xp))

#Find columns in xp where the name contains 'Treatment'
    c2 <- grepl('Treatment', colnames(xp))

#Find rows in pdat that correspond to either 'Control' or 'Treatment'
    r1 <- with(pdat, Group == 'Control')
    r2 <- with(pdat, Group == 'Treatment')

# In xp: find the rows that correspond to Control or Treatment, those that do not match will be
    #set to zero. Then, substract the values from the rows corresponding to 'Control' from those that correspond
    #to 'Treatment'
    X <- xp[r1, ] - xp[r2, ]

    ## remove columns that do not contain name 'Control' or 'Treatment'
    X[, ! (c1 | c2)] <- 0
    ## zero out the parametric cols, those that do not contain in the characters 's('
    #X[, !grepl('^s\\(', colnames(xp))] <- 0

    #Multiply matrix by model coefficients. X has (p,n) (rows, columns) and the coefficient matrix has
    #dimensions (n,1). The resulting matrix has dimensions (p,1)
    dif <- X %*% coef(m1)

    #comp<-test %*% coef(gam1)[3:10]

#Calculate standard error for the computed differences using the variance-covariance matrix
    #of the model
    se <- sqrt(rowSums((X %*% vcov(m1, unconditional = FALSE)) * X))
    crit <- qt(0.05/2, df.residual(m1), lower.tail = FALSE)
    #upper  limits
    upr <- dif + (crit * se)
    #lower limits
    lwr <- dif - (crit * se)
    #put all components in a dataframe for plotting
    comp1<-data.frame(pair = paste('Control', 'Treatment', sep = '-'),
               diff = dif,
               se = se,
               upper = upr,
               lower = lwr)

#add time point sequence
comp_StO2 <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp1))

#use function from the pairwise comparison plot in the manuscript to get the shaded regions

    my_list<-pairwise_limits(comp_StO2)
  rib_col<-'#EDD03AFF' #color for the ribbon
#plot the difference
c1<-ggplot(comp_StO2, aes(x = Day, y = diff, group = pair)) +
  #shaded region
  annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-10,
             label="Control",size=10
           )+
  #shaded region  
  annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5
           ) +
  annotate("text",
             x=6,
             y=-10,
             label="Treatment",
             size=10
           )+
  #ribbon for difference confidence interval  
  geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(color='black',size=1) +
    geom_line(data=comp_StO2,aes(y=0),size=0.5)+
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )
\end{verbatim}

(ref:pairwise-comp-caption) Smooth pairwise comparisons for model
\texttt{m1} using a 95\% confidence interval for the difference between
smooths. The comparison is centered at the response scale. Shaded
regions indicate time intervals where each treatment group has a
non-zero effect.

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/pairwise-comp-workflow-fig-1}

}

\caption{(ref:pairwise-comp-caption)}\label{fig:pairwise-comp-workflow-fig}
\end{figure}

Of notice, a convenient wrapper for the function described above exists
in the package \texttt{gratia}. In this package,
\texttt{difference\_smooths} is a function that makes the comparisons
and produces Figure @ref(fig:pairwise-comp-workflow-fig) when is used on
a fitted model. The function syntax and an example can be found at:

\url{https://cran.r-project.org/web/packages/gratia/gratia.pdf}

Keep in mind that this function \textbf{does not} center the pairwise
comparison at the response scale, so it has to be shifted in order to be
compared to the raw data.

\hypertarget{gam-and-linear-model-plots-and-missing-data}{%
\section{GAM and Linear model plots and Missing
data}\label{gam-and-linear-model-plots-and-missing-data}}

This section covers the code used to generate Figure
@ref(fig:sim-smooth-plot), where the simulated data, fit of the
``final'' GAM (\texttt{m1}), linear model and GAM on data with missing
observations are presented. Note that panel A in Figure
@ref(fig:sim-smooth-plot) and the inset are generated in the code chunk
where the data is simulated in Section @ref(tumor-data-simulation), and
are called later to build the figure.

\hypertarget{gam-and-linear-model-plots}{%
\subsection{GAM and Linear model
plots}\label{gam-and-linear-model-plots}}

This code chunk creates panels B and D in Figure
@ref(fig:sim-smooth-plot). Note that this code uses the final GAM from
the previous section (\texttt{m1}), so the simulated data and the model
should be generated before running this section.

\begin{verbatim}
#linear model
lm1<-lm(StO2_sim ~ Day + Group + Day * Group, data = dat_sim)

#creates a dataframe using the length of the covariates for the GAM
gam_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1),
                         subject=factor(rep(1:10)))

#creates a dataframe using the length of the covariates for rm-ANOVA
lm_predict<-expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = c(0:10),
                        subject=factor(rep(1:10)),
                          )
lm_predict$subject<-factor(paste(lm_predict$subject, lm_predict$Group, sep = "-"))

#adds the predictions to the grid and creates a confidence interval for GAM
gam_predict<-gam_predict%>%
    mutate(fit = predict(m1,gam_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(m1, gam_predict,se.fit = TRUE,type='response')$se.fit)

#using lm
lm_predict<-lm_predict%>%
    mutate(fit = predict(lm1,lm_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(lm1, lm_predict,se.fit = TRUE,type='response')$se.fit)

#plot smooths and confidence interval for GAM
f3<-ggplot(data=dat_sim, aes(x=Day, y=StO2_sim, group=Group)) +
    geom_point(aes(color=Group),size=1.5,alpha=0.5,show.legend = FALSE)+
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),
                   fill=Group
                   ),
              alpha=0.3,
              data=gam_predict,
            show.legend=FALSE,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=gam_predict,
              show.legend = FALSE)+
  #facet_wrap(~Group)+
  labs(y=expression(atop(StO[2],'complete')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
  theme(
    axis.text=element_text(size=22)
  )+
      thm+
  thm1

#plot linear fit for rm-ANOVA
f4<-ggplot(data=dat_sim, aes(x=Day, y=StO2_sim, group=Group)) +
    geom_point(aes(color=Group),size=1.5,alpha=0.5,show.legend = FALSE)+
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),fill=Group),
              alpha=0.3,
              data=lm_predict,
              show.legend = FALSE,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=lm_predict,
              show.legend = FALSE)+
  #facet_wrap(~Group)+
  labs(y=expression(paste('StO'[2],' (simulated)')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
  theme(
    axis.text=element_text(size=22)
  )+
      thm+
  thm1

#posthoc comparisons for the linear model
#library(multcomp)

#summary(glht(lm1, linfct = mcp(Group = 'Tukey')))
#summary(glht(lm1, linfct=mcp(Group="Tukey", interaction_average=TRUE)))
\end{verbatim}

\hypertarget{working-with-missing-data-in-gams}{%
\subsection{Working with Missing data in
GAMs}\label{working-with-missing-data-in-gams}}

This code chunk first randomly deletes 40\% of the total observations in
the original simulated data, and then an interaction GAM is fitted to
the remaining data. Model diagnostics are presented, and an object that
stores the fitted smooths is saved to be called in the final code chunk
to build the figure.

\begin{verbatim}
#missing data
#create a sequence of 40 random numbers between 1 and 100, these numbers will
#correspond to the row numbers to be randomly erased from the original dataset

missing <- sample(1:100, 40)

#create a new dataframe from the simulated data with 40 rows randomly removed, keep the missing values as NA

ind <- which(dat_sim$StO2_sim %in% sample(dat_sim$StO2_sim, 40))

#create a new dataframe, remove the StO2 column
dat_missing <- dat_sim[,-1]

#add NAs at the ind positions
dat_missing$StO2_sim[ind]<-NA

#Count the number of remaining observations per day (original dataset had 10 per group per day)
dat_missing %>%
    group_by(Day,Group) %>%
    filter(!is.na(StO2_sim))%>%
  count(Day)

#the same model used for the full dataset
mod_m1 <- gam(StO2_sim ~ Group+s(Day,by=Group,k=5), data  = dat_missing,family=scat)
#appraise the model
appraise(mod_m1)

m_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1))

#adds the predictions to the grid and creates a confidence interval
m_predict<-m_predict%>%
    mutate(fit = predict(mod_m1,m_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(mod_m1, m_predict,se.fit = TRUE,type='response')$se.fit)

f6<-ggplot(data=dat_missing, aes(x=Day, y=StO2_sim, group=Group)) +
    geom_point(aes(color=Group),size=1.5,alpha=0.5,show.legend = FALSE)+
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),
                   fill=Group
                   ),
              alpha=0.3,
              data=m_predict,
            show.legend=FALSE,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=m_predict,
              show.legend = TRUE)+
  #facet_wrap(~Group)+
  labs(y=expression(atop(StO[2],'missing')))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
      theme_classic()+
  theme(
    axis.text=element_text(size=22)
  )+
      thm+
  thm1
\end{verbatim}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/sim-smooth-plot-Appendix-1}

}

\caption{(ref:sim-plot-caption)}\label{fig:sim-smooth-plot-Appendix}
\end{figure}

\hypertarget{pairwise-comparisons-in-gams-full-and-missing-data-cases}{%
\subsection{Pairwise comparisons in GAMs: full and missing data
cases}\label{pairwise-comparisons-in-gams-full-and-missing-data-cases}}

The next code chunk reproduces Figure @ref(fig:plot-pairwise-comp). Here
pairwise comparisons are made for the full and missing datasets.

\begin{verbatim}
##Pairwise comparisons

pdat <- expand.grid(Day = seq(0, 10, length = 400),
                    Group = c('Control', 'Treatment'))

#this function takes the model, grid and groups to be compared using the lpmatrix
#originally developed by G. Simpson:
#https://fromthebottomoftheheap.net/2017/10/10/difference-splines-i/

smooth_diff <- function(model, newdata, g1, g2, alpha = 0.05,
                        unconditional = FALSE) {
    xp <- predict(model, newdata = newdata, type = 'lpmatrix')
    #Find columns in xp where the name contains "Control" and "Treatment"
    col1 <- grepl(g1, colnames(xp))
    col2 <- grepl(g2, colnames(xp))
    #Find rows in xp that correspond to each treatment
    row1 <- with(newdata, Group == g1)
    row2 <- with(newdata, Group == g2)
    ## difference rows of xp for data from comparison
    X <- xp[row1, ] - xp[row2, ]
    ## zero out cols of X related to splines for other lochs
    X[, ! (col1 | col2)] <- 0

    ## zero out the parametric cols
    #This line has been commented to keep the comparison at the response level,
    #otherwise it gives the marginal change between smooths
    #X[, !grepl('^s\\(', colnames(xp))] <- 0
    dif <- X %*% coef(model)
    #get standard error, critical value and boundaries
    se <- sqrt(rowSums((X %*% vcov(model, unconditional = unconditional)) * X))
    crit <- qt(alpha/2, df.residual(model), lower.tail = FALSE)
    upr <- dif + (crit * se)
    lwr <- dif - (crit * se)
    data.frame(pair = paste(g1, g2, sep = '-'),
               diff = dif,
               se = se,
               upper = upr,
               lower = lwr)
}

#use the function to calculate the difference in smooths
comp1<-smooth_diff(m1,pdat,'Control','Treatment')

#Create a dataframe with time, comparisons and labels for regions where difference exists
comp_StO2_full <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp1)) %>%
  mutate(interval=case_when(
    upper>0 & lower<0~"no-diff",
    upper<0~"less",
    lower>0~"greater"
  ))

pairwise_limits<-function(dataframe){
    #extract values where the lower limit of the ribbon is greater than zero
    #this is the region where the control group effect is greater
    v1<-dataframe%>%
        filter(lower>0)%>%
        select(Day)
    #get day  initial value
    init1=v1$Day[[1]]
    #get day final value
    final1=v1$Day[[nrow(v1)]]

    #extract values where the value of the upper limit of the ribbon is lower than zero
    #this corresponds to the region where the treatment group effect is greater
    v2<-comp_StO2_full%>%
        filter(upper<0)%>%
        select(Day)

    init2=v2$Day[[1]]
    final2=v2$Day[[nrow(v2)]]
    #store values
    my_list<-list(init1=init1,
                  final1=final1,
                  init2=init2,
                  final2=final2)
return(my_list)
}

my_list<-pairwise_limits(comp_StO2_full)
rib_col<-'#EDD03AFF'

c1<-ggplot(comp_StO2_full, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
                xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
                fill='#30123BFF',
                alpha = 0.5,
                ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8,
           angle=90
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8,
           angle=90
           )+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(data=comp_StO2_full,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +

    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

###for missing data
comp2<-smooth_diff(mod_m1,pdat,'Control','Treatment')
comp_StO2_missing <- cbind(Day = seq(0, 10, length = 400),
                   rbind(comp2))

missing_plot<-ggplot(comp_StO2_missing, aes(x = Day, y = diff, group = pair)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2) +
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    labs(x = 'Days', 
         y = expression(paste('Difference in StO'[2],'\n (missing data)' 
                              )))+
  scale_x_continuous(breaks=c(0,2,5,7,10))+
  theme_classic()+
  theme(
     text=element_text(size=18),
     legend.title=element_blank()
     )

my_list<-pairwise_limits(comp_StO2_missing)

c2<-ggplot(comp_StO2_missing, aes(x = Day, y = diff, group = pair)) +
    annotate("rect",
             xmin =my_list$init1, xmax =my_list$final1,ymin=-Inf,ymax=Inf,
             fill='#30123BFF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=1.5,
             y=-18,
             label="Control>Treatment",
           size=8
           )+
    annotate("rect",
             xmin =my_list$init2, xmax =my_list$final2,ymin=-Inf,ymax=Inf,
             fill='#7A0403FF',
             alpha = 0.5,
    ) +
  annotate("text",
             x=6,
             y=-18,
             label="Treatment>Control",
             size=8)+
    geom_ribbon(aes(ymin = lower, ymax = upper),
                alpha = 0.5,
                fill=rib_col) +
    geom_line(data=comp_StO2_missing,aes(y=0),size=0.5)+
    geom_line(color='black',size=1) +
    facet_wrap(~ pair) +
    theme_classic()+
    labs(x = 'Days', y = expression(paste('Difference in StO'[2] )))+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    theme(
        text=element_text(size=18),
        legend.title=element_blank()
    )

pair_comp<-c1+c2
\end{verbatim}

\begin{figure}

{\centering \includegraphics[width=0.75\linewidth]{Full_document_SAGE_files/figure-latex/plot-pairwise-comp-Appendix-1}

}

\caption{(ref:pairwise-comp)}\label{fig:plot-pairwise-comp-Appendix}
\end{figure}

\bibliographystyle{sagev}
\bibliography{refs1.bib}

\end{document}