### 2021-11-03 Manuscripts/Manuscript_by_chapters-SIM_Revisions/response_to_reviewers/Response_to_reviewers.Rmd
-Working on reviewers' comments -Reorganized folders

---
title: |
    | **Generalized additive models to analyze biomedical non-linear longitudinal data in R:**
    | Beyond repeated measures ANOVA and Linear Mixed Models
    | 
    | Response to reviewers
header-includes:
    \usepackage{placeins}
output:  
  bookdown::pdf_document2:
    #template: my-template.tex #if a custom template that removes the additional "and" in the author information is desired
    pandoc_args: --listings #calls the listings package to fit code within the page margins
    keep_tex: yes #keep LaTeX file for submission
    fig_caption: yes #allows captions in figures
    toc: false #do not include table of contents
    extra_dependencies:
      subfig: null #allows for subfigures
      breqn: null #line breaks for long equations
      caption: ["font={small}"] #size of the figure captions
      #float: null #allows for control of placement of figures
    includes:
      in_header: latex_docs/preamble_Appendix.sty #additional LaTeX formatting
  bookdown::word_document2:
    fig_caption: yes #figure caption
    keep_md: yes
  bookdown::html_document2:
    css: "style.css" #style for the HTML document
csl: elsevier-with-titles.csl #style for references
bibliography: refs.bib #references
link-citations: yes #adds links to the citations
'': default
---

# General Comments

# Comments from reviewers
## Missing data

The title for Section 3.4 has been changed from "Missing observations" to "Unbalanced data". We have restructured this section in order to:

- Avoid confusion by using the term "missing data". We now refer to different number of observations as "unbalanced data" (L201-204).

- Make clear that LMEMs also can work with missing observations (L209)

- Emphasize that GAMs are not immune to missing data problems (L217-220)

## Unnecesary restriction of LMMs

## GLMMs

## GLMs, GAMs, and conditional distributions

## Thin plate regression splines

The following changes have been made:

- L297-300 no longer refer to "knots" when referring to the construction of the smooth, now indicating that the number of _basis functions_ is what is specified when working with thin plate regression splines (TPRS), in the updated manuscript this appears in L310-315.

## Penalised splines

## Bayesian

## Coverage of confidence intervals

## Differences in smooths

## Appendix