### 2021-11-04 Manuscripts/Manuscript_by_chapters-SIM_Revisions/response_to_reviewers/Response_to_reviewers.Rmd
Updated manuscript and response to reviewers

---
title: |
    | **Generalized additive models to analyze biomedical non-linear longitudinal data in R:**
    | Beyond repeated measures ANOVA and Linear Mixed Models
    | 
    | Response to reviewers
header-includes:
    \usepackage{placeins}
output:  
  bookdown::pdf_document2:
    #template: my-template.tex #if a custom template that removes the additional "and" in the author information is desired
    pandoc_args: --listings #calls the listings package to fit code within the page margins
    keep_tex: yes #keep LaTeX file for submission
    fig_caption: yes #allows captions in figures
    toc: false #do not include table of contents
    extra_dependencies:
      subfig: null #allows for subfigures
      breqn: null #line breaks for long equations
      caption: ["font={small}"] #size of the figure captions
      #float: null #allows for control of placement of figures
    includes:
      in_header: latex_docs/preamble_Appendix.sty #additional LaTeX formatting
  bookdown::word_document2:
    fig_caption: yes #figure caption
    keep_md: yes
  bookdown::html_document2:
    css: "style.css" #style for the HTML document
#csl: elsevier-with-titles.csl #style for references
#bibliography: refs.bib #references
link-citations: yes #adds links to the citations
'': default
---

# General Comments

# Comments from reviewers
## Missing data

The title for Section 3.4 has been changed from "Missing observations" to "Unbalanced data". We have restructured this section in order to:

- Avoid confusion by using the term "missing data". We now refer to different number of observations as "unbalanced data" (L201-204).

- Make clear that LMEMs also can work with missing observations (L209)

- Emphasize that GAMs are not immune to missing data problems (L217-220)

## Unnecesary restriction of LMMs

## GLMMs

## GLMs, GAMs, and conditional distributions

## Thin plate regression splines

The following changes have been made:

- The information on L289-294 pertaining splines has been changed. L296-307 now give a better description of cubic splines (CS), indicating their limitation due to knot placement. The lack of knot placement for thin plate regression splines (TPRS) is indicated as an advantage over CS.

- L297-300 no longer refer to "knots" when referring to the construction of the smooth, now indicating that the number of _basis functions_ is what is specified when working with TPRS. In the updated manuscript this appears in L310-315.

## Penalised splines

## Bayesian

- We have clarified the concept "empirical Bayesian", which appeared on L318-L319 in the original manuscript. In the revised manuscript, L335-336 now indicate that Stan, JAGS or other probabilistic programming language can be used to estimate GAMs using a full Bayesian approach.

- The sentence "Moreover, the use of the restricted maximum likelihood (REML) to estimate the smoothing parameters gives an empirical estimate of the smooth model", which appears in L319 in the original manuscript has been removed from the text. Instead, the concept of REML has been moved to Section 6.2 L390-393, where we state some of the reasons indicated by Wood when choosing restricted maximum likelihood (REML) over the default general cross validation (GCV) method for smooth parameter estimation in _mgcv_.

## Coverage of confidence intervals

The content in L320-329 regarding confidence intervals (CIs) in the original submission has been changed in the following manner:

- L339-L352 now contain a more detailed and accurate explanation on the differences and interpretation of "pointwise" CIs and "across-the-function" CIs. We also reference the work of Marra and Wood if the reader desires a more in-depth exploration.

## Differences in smooths

## Appendix