### 2021-12-26 Code and miscellaneous/Interval_comparisons.Rmd

---
title: Testing the simultaneous CI functions
---

```{r setup, message = FALSE,include=FALSE}
#the packages needed to run the code

library(patchwork)
library(tidyverse)
library(mvnfast)
library(nlme)
library(mgcv)
library(gratia)
library(scico)
set.seed(2021) #set seed for reproducibility
thm1<-scale_fill_scico_d(palette="tokyo",begin=0.3, end=0.8, direction = -1, aesthetics = c("colour","fill"))
```

I want to test if there's a difference because when knitting I get the message:

`In gratia:::get_vcov(model, unconditional = unconditional, frequentist = frequentist) :Covariance corrected for smoothness uncertainty not available.`

```{r,generate-simulated-data}
#data, "the truth"
dat<-tibble(StO2=c(4,27,3,2,0.5,7,4,50,45,56),
            Day=rep(c(0,2,5,7,10),times=2),
            Group=as.factor(rep(c("Control","Treatment"),each=5))
)

#simulate data
simulate_data <- function(dat, n = 10, sd = 5) {
    dat_sim <- dat %>%
        slice(rep(1:n(), each = n)) %>%
        group_by(Group, Day) %>%
        mutate(
            StO2_sim = pmax(rnorm(n, StO2, sd), 0.0001),
            subject=rep(1:10),
            subject=factor(paste(subject, Group, sep = "-"))
        ) %>%
        ungroup()

    return(dat_sim)
}

#subject = factor(paste(subject, treatment, sep = "-")))

n <- 10 #number of observations
sd <- 10 #approximate sd from paper
df <- 6
dat_sim <- simulate_data(dat, n, sd)

#plot data

c<-ggplot(dat_sim, aes(x = Day, 
                        y = StO2_sim, 
                        color = Group,
                        group=subject)) +

    geom_line(size=0.6, alpha=0.4,show.legend=FALSE)+
    geom_line(aes(x = Day, 
               y = StO2,
               color=Group),
           size=1.5,
           data=dat,
           inherit.aes=FALSE,
              show.legend = FALSE)+
  labs(y=expression(atop(StO[2],'(simulated)')))+
  theme_classic()+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    thm1

```

Now, I fit the GAM.

```{r,fit-GAM}
#GAM, interaction of time and treatment

gam_02 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5),
              method='REML',
              data  = dat_sim)
```

Now I follow the exact code from G. Simpson's post to generate the simultaneous interval (link to the original post below):
https://fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/

```{r,simulatenous-interval}
#Generate random values from multivariate normal
rmvn <- function(n, mu, sig) { ## MVN random deviates
    L <- mroot(sig)
    m <- ncol(L)
    t(mu + L %*% matrix(rnorm(m*n), m, n))
}

#Bayesian covariance matrix of the model coefficients
Vb <- vcov(gam_02)

#grid of values for confidence band
newd <- expand_grid(Group = factor(c("Control", "Treatment")),
                                   Day = seq(0, 10, by = 0.1),
                                   subject=factor(rep(1:10)))

#Generate predictions and standard errors
pred <- predict(gam_02, newd, se.fit = TRUE)
se.fit <- pred$se.fit

#number of draws
N <- 10000

#bias of the estimated model coefficients
BUdiff <- rmvn(N, mu = rep(0, nrow(Vb)), sig = Vb)

#evaluate the basis functions at the locations of x
Cg <- predict(gam_02, newd, type = "lpmatrix")

#Computes deviations between the fitted and true parameters
simDev <- Cg %*% t(BUdiff)

#compute absolute values of the standardized deviations from the true model
absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))

#find maximum of the absolute standardized deviations at the grid of values of x for each simulation
masd <- apply(absDev, 2L, max)

#Find critical value used to scale the standard errors to yield the simultaneous interval

crit <- quantile(masd, prob = 0.95, type = 8)

pred <- transform(cbind(data.frame(pred), newd),
                  uprP = fit + (2 * se.fit),
                  lwrP = fit - (2 * se.fit),
                  uprS = fit + (crit * se.fit),
                  lwrS = fit - (crit * se.fit))

#plot the simultaneous and pointwise intervals
a<-ggplot(pred, aes(x =Day,group=Group,color=Group,fill=Group)) +
    geom_path(aes(y=fit),lwd=2,show.legend = FALSE)+
    geom_ribbon(aes(ymin = lwrS, ymax = uprS), alpha = 0.2,show.legend = FALSE) +
    geom_ribbon(aes(ymin = lwrP, ymax = uprP), alpha = 0.2,show.legend = FALSE) +
    labs(y = "G. Simpson's code")+
   geom_hline(yintercept = 60,color="red")+
    geom_hline(yintercept = 20,color="red")+
    geom_hline(yintercept = 0,color="red")+
    theme_classic()+
    thm1

```

Now I will use the code from our paper to get the same intervals

```{r,code-from-paper}

ci <- confint(gam_02, parm = "s(Day)", partial_match = TRUE, type = "confidence")
## simultaneous interval
si <- confint(gam_02, parm = "s(Day)", type = "simultaneous", partial_match = TRUE)

# mean shift for group 2

Intercept <-coef(gam_02)[1]
const <- coef(gam_02)[2]+Intercept

ci <- ci %>%
mutate(est = case_when(Group == "Treatment" ~ est + const,
TRUE ~ est+Intercept),
lower = case_when(Group == "Treatment" ~ lower + const,
TRUE ~ lower+Intercept),
upper = case_when(Group == "Treatment" ~ upper + const,
TRUE ~ upper+Intercept))

si <- si %>%
mutate(est = case_when(Group == "Treatment" ~ est + const,
TRUE ~ est+Intercept),
lower = case_when(Group == "Treatment" ~ lower + const,
TRUE ~ lower+Intercept),
upper = case_when(Group == "Treatment" ~ upper + const,
TRUE ~ upper+Intercept))

b<-ggplot(ci, aes(x = Day, y = est, group = smooth)) +
geom_line(lwd = 1,show.legend = FALSE) +
geom_ribbon(data = ci, mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill = Group),
inherit.aes = FALSE, alpha = 0.7,
show.legend=FALSE) +
geom_ribbon(data = si,
mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill =Group),
inherit.aes = FALSE, alpha = 0.4,
show.legend=FALSE)+
    geom_line(data=si,aes(Day,upper,color=Group),show.legend = FALSE)+
    geom_line(data=si,aes(Day,lower,color=Group),show.legend = FALSE)+
    geom_hline(yintercept = 60,color="red")+
    geom_hline(yintercept = 20,color="red")+
    geom_hline(yintercept = 0,color="red")+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
     labs(y = "Our code")+
      theme_classic()+
        thm1

```

Now I plot both results and is strange, but the intervals that we get from G. Simpson's code are wider! why?

```{r,plotting-data-and-intervals}
#plot both
c+a+b+scale_y_continuous(breaks=c(-20,0,20,40,60))

```

```{r,poitwise-diff}
##this function determines the pointwise confidence interval of a difference between two smooths

difference_pointwise <- function(f1, f2, smooth, by_var, smooth_var, data, Xp, V, coefs, nrep = 1000) {
    ## make sure f1 and f2 are characters
    f1 <-  as.character(f1)
    f2 <-  as.character(f2)
    cnames <- colnames(Xp)
    ## columns of Xp associated with pair of smooths
    c1 <- grepl(gratia:::mgcv_by_smooth_labels(smooth, by_var, f1), cnames, fixed = TRUE)
    c2 <- grepl(gratia:::mgcv_by_smooth_labels(smooth, by_var, f2), cnames, fixed = TRUE)
    ## rows of Xp associated with pair of smooths
    r1 <- data[[by_var]] == f1
    r2 <- data[[by_var]] == f2

    ## difference rows of Xp for pair of smooths
    X <- Xp[r1, ] - Xp[r2, ]

    ######IMPORTANT: uncommenting the following two lines
    #removes the group means from the comparison######

    ## zero the cols related to other splines
    # X[, ! (c1 | c2)] <- 0

    ## zero out the parametric cols
    #X[, !grepl('^s\\(', cnames)] <- 0

    ## compute difference
    sm_diff <- drop(X %*% coefs)
    se <- sqrt(rowSums((X %*% V) * X))
    nr <- NROW(X)

    ## Calculate posterior simulation for smooths
    coefs_sim <- t(rmvn(nrep, rep(0, nrow(V)), V))
    rownames(coefs_sim) <- rownames(V)
    simDev <- X %*% coefs_sim
    absDev <- abs(sweep(simDev, 1, se, FUN = "/"))
    masd <- apply(absDev, 2, max)
    crit_s <- quantile(masd, prob = 0.95, type = 8)

    out <- list(smooth = rep(smooth, nr), by = rep(by_var, nr),
                level_1 = rep(f1, nr),
                level_2 = rep(f2, nr),
                diff = sm_diff, se = se,
                lower_s = sm_diff - crit_s * se,
                upper_s = sm_diff + crit_s*se)

    out <- new_tibble(out, nrow = NROW(X), class = "difference_smooth")
    ## Only need rows associated with one of the levels
    out <- bind_cols(out, data[r1, smooth_var])

    out
}

```

And this:

```{r,difference-smooths}
#this function calculates the pointwise (by calling pointwise_comparisons.R) CI and the simultaneous CI
# for a pairwise comparison between two smooths
difference_smooths <- function(model,
                               smooth,
                               n = 100,
                               ci_level = 0.95,
                               newdata = NULL,
                               partial_match = TRUE,
                               unconditional = FALSE,
                               frequentist = FALSE,
                               nrep = 10000,
                               include_means = TRUE,
                               ...) {
    if (missing(smooth)) {
        stop("Must specify a smooth to difference via 'smooth'.")
    }

    # smooths in model
    S <- gratia::smooths(model) # vector of smooth labels - "s(x)"
    # select smooths
    select <-
        gratia:::check_user_select_smooths(smooths = S, select = smooth,
                                           partial_match = partial_match)#,
    # model_name = expr_label(substitute(object)))
    sm_ids <- which(select)
    smooths <- gratia::get_smooths_by_id(model, sm_ids)
    sm_data <- map(sm_ids, gratia:::smooth_data,
                   model = model, n = n, include_all = TRUE)
    sm_data <- bind_rows(sm_data)
    by_var <- by_variable(smooths[[1L]])
    smooth_var <- gratia:::smooth_variable(smooths[[1L]])
    pairs <- as_tibble(as.data.frame(t(combn(levels(sm_data[[by_var]]), 2)),
                                     stringsAsFactor = FALSE))
    names(pairs) <- paste0("f", 1:2)

    Xp <- predict(model, newdata = sm_data, type = "lpmatrix")
    V <- gratia:::get_vcov(model, unconditional = unconditional,
                           frequentist = frequentist)
    coefs <- coef(model)

    out <- pmap(pairs, difference_pointwise, smooth = smooth, by_var = by_var,
                smooth_var = smooth_var, data = sm_data, Xp = Xp, V = V,
                coefs = coefs, nrep = nrep)
    out <- bind_rows(out)
    crit <- qnorm((1 - ci_level) / 2, lower.tail = FALSE)

    out <- add_column(out,
                      lower = out$diff - (crit * out$se),
                      upper = out$diff + (crit * out$se),
                      .after = 6L)
    out
}
```

Let's get the fit then:

```{r,testing}
diff_testing <- difference_smooths(gam_02, smooth = "s(Day)", newdata = newdat,
                     unconditional = TRUE, frequentist = FALSE,
                     n=100, partial_match = TRUE, nrep=10000)

```