### 2021-12-27 Code and miscellaneous/Interval_comparisons.Rmd
Checking code as there is something going on with the CI computation

---
title: Testing the simultaneous CI functions
---

```{r setup, message = FALSE,include=FALSE}
#the packages needed to run the code

library(patchwork)
library(tidyverse)
library(mvnfast)
library(nlme)
library(mgcv)
library(gratia)
library(scico)
set.seed(2021) #set seed for reproducibility
thm1<-scale_fill_scico_d(palette="tokyo",begin=0.3, end=0.8, direction = -1, aesthetics = c("colour","fill"))
```

I want to test if there's a difference because when knitting I get the message:

`In gratia:::get_vcov(model, unconditional = unconditional, frequentist = frequentist) :Covariance corrected for smoothness uncertainty not available.`

```{r,generate-simulated-data}
#data, "the truth"
dat<-tibble(StO2=c(4,27,3,2,0.5,7,4,50,45,56),
            Day=rep(c(0,2,5,7,10),times=2),
            Group=as.factor(rep(c("Control","Treatment"),each=5))
)

#simulate data
simulate_data <- function(dat, n = 10, sd = 5) {
    dat_sim <- dat %>%
        slice(rep(1:n(), each = n)) %>%
        group_by(Group, Day) %>%
        mutate(
            StO2_sim = pmax(rnorm(n, StO2, sd), 0.0001),
            subject=rep(1:10),
            subject=factor(paste(subject, Group, sep = "-"))
        ) %>%
        ungroup()

    return(dat_sim)
}

#subject = factor(paste(subject, treatment, sep = "-")))

n <- 10 #number of observations
sd <- 10 #approximate sd from paper
df <- 6
dat_sim <- simulate_data(dat, n, sd)

#plot data

c<-ggplot(dat_sim, aes(x = Day, 
                        y = StO2_sim, 
                        color = Group,
                        group=subject)) +

    geom_line(size=0.6, alpha=0.4,show.legend=FALSE)+
    geom_line(aes(x = Day, 
               y = StO2,
               color=Group),
           size=1.5,
           data=dat,
           inherit.aes=FALSE,
              show.legend = FALSE)+
  labs(y=expression(atop(StO[2],'(simulated)')))+
  theme_classic()+
    scale_x_continuous(breaks=c(0,2,5,7,10))+
    thm1

```

Now, I fit the GAM.

```{r,fit-GAM}
#GAM, interaction of time and treatment

gam_02 <- gam(StO2_sim ~ Group+s(Day, by = Group, k = 5),
              method='REML',
              data  = dat_sim)
```

Now I follow the exact code from G. Simpson's post to generate the simultaneous interval (link to the original post below):
https://fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/

```{r,simulatenous-interval}
#Generate random values from multivariate normal
rmvn <- function(n, mu, sig) { ## MVN random deviates
    L <- mroot(sig)
    m <- ncol(L)
    t(mu + L %*% matrix(rnorm(m*n), m, n))
}

#Bayesian covariance matrix of the model coefficients
Vb <- vcov(gam_02)

#grid of values for confidence band
newd <- expand_grid(Group = factor(c("Control", "Treatment")),
                                   Day = seq(0, 10, by = 0.1),
                                   subject=factor(rep(1:10)))

#Generate predictions and standard errors
pred <- predict(gam_02, newd, se.fit = TRUE)
se.fit <- pred$se.fit

#number of draws
N <- 10000

#bias of the estimated model coefficients
BUdiff <- rmvn(N, mu = rep(0, nrow(Vb)), sig = Vb)

#evaluate the basis functions at the locations of x
Cg <- predict(gam_02, newd, type = "lpmatrix")

#Computes deviations between the fitted and true parameters
simDev <- Cg %*% t(BUdiff)

#compute absolute values of the standardized deviations from the true model
absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))

#find maximum of the absolute standardized deviations at the grid of values of x for each simulation
masd <- apply(absDev, 2L, max)

#Find critical value used to scale the standard errors to yield the simultaneous interval

crit <- quantile(masd, prob = 0.95, type = 8)

pred <- transform(cbind(data.frame(pred), newd),
                  uprP = fit + (2 * se.fit),
                  lwrP = fit - (2 * se.fit),
                  uprS = fit + (crit * se.fit),
                  lwrS = fit - (crit * se.fit))

#plot the simultaneous and pointwise intervals
a<-ggplot(pred, aes(x =Day,group=Group,color=Group,fill=Group)) +
    geom_path(aes(y=fit),lwd=2,show.legend = FALSE)+
    geom_ribbon(aes(ymin = lwrS, ymax = uprS), alpha = 0.4,show.legend = FALSE) +
    geom_ribbon(aes(ymin = lwrP, ymax = uprP), alpha = 0.7,show.legend = FALSE) +
    labs(y = "G. Simpson's code")+
   geom_hline(yintercept = 60,color="red")+
    geom_hline(yintercept = 35.77,color="red")+
    geom_hline(yintercept = 20,color="red")+
    geom_hline(yintercept = 0,color="red")+
    scale_x_continuous(breaks=seq(1,10,1))+
    theme_classic()+
    thm1

```

Now I will use the code from our paper to get the same intervals

```{r,code-from-paper}

ci <- confint(gam_02, parm = "s(Day)", partial_match = TRUE, type ="confidence")
## simultaneous interval
si <- confint(gam_02, parm = "s(Day)", type = "simultaneous", partial_match = TRUE,nsim=10000)

# mean shift for group 2

Intercept <-coef(gam_02)[1]
const <- coef(gam_02)[2]+Intercept

ci <- ci %>%
mutate(est = case_when(Group == "Treatment" ~ est + const,
TRUE ~ est+Intercept),
lower = case_when(Group == "Treatment" ~ lower + const,
TRUE ~ lower+Intercept),
upper = case_when(Group == "Treatment" ~ upper + const,
TRUE ~ upper+Intercept))

si <- si %>%
mutate(est = case_when(Group == "Treatment" ~ est + const,
TRUE ~ est+Intercept),
lower = case_when(Group == "Treatment" ~ lower + const,
TRUE ~ lower+Intercept),
upper = case_when(Group == "Treatment" ~ upper + const,
TRUE ~ upper+Intercept))

b<-ggplot(ci, aes(x = Day, y = est, group = smooth)) +
geom_line(lwd = 1,show.legend = FALSE) +
geom_ribbon(data = ci, mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill = Group),
inherit.aes = FALSE, alpha = 0.7,
show.legend=FALSE) +
geom_ribbon(data = si,
mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill =Group),
inherit.aes = FALSE, alpha = 0.4,
show.legend=FALSE)+
    geom_line(data=si,aes(Day,upper,color=Group),show.legend = FALSE)+
    geom_line(data=si,aes(Day,lower,color=Group),show.legend = FALSE)+
    geom_hline(yintercept = 60,color="red")+
    geom_hline(yintercept = 20,color="red")+
    geom_hline(yintercept = 0,color="red")+
    geom_hline(yintercept = 35.77,color="red")+
    scale_x_continuous(breaks=seq(1,10,1))+
     labs(y = "Our code")+
      theme_classic()+
        thm1

```

Additional test: use the approach we used initially to draw the confidence intervals, see if there's a difference:

```{r}

#creates a dataframe using the length of the covariates for the GAM
gam_predict <- expand_grid(Group = factor(c("Control", "Treatment")),
                         Day = seq(0, 10, by = 0.1),
                         subject=factor(rep(1:10)))

#adds the predictions to the grid and creates a confidence interval for GAM
gam_predict<-gam_predict%>%
    mutate(fit = predict(gam_02,gam_predict,se.fit = TRUE,type='response')$fit,
           se.fit = predict(gam_02, gam_predict,se.fit = TRUE,type='response')$se.fit)

#plot smooths and confidence interval for GAM
f3<-ggplot(data=dat_sim, aes(x=Day, y=StO2_sim, group=Group)) +
  geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),
                   fill=Group
                   ),
              alpha=0.3,
              data=gam_predict,
            show.legend=FALSE,
                inherit.aes=FALSE) +
  geom_line(aes(y=fit,
                color=Group),
              size=1,data=gam_predict,
              show.legend = FALSE)+
    geom_hline(yintercept = 60,color="red")+
    geom_hline(yintercept = 20,color="red")+
    geom_hline(yintercept = 0,color="red")+
    geom_hline(yintercept = 35.77,color="red")+
  labs(y='previous method')+
    scale_x_continuous(breaks=seq(1,10,1))+
      theme_classic()+
    thm1

```

Now I plot both results and is strange, but the intervals that we get from G. Simpson's code are wider! why?

```{r,plotting-data-and-intervals}
#plot both
c+a+b+f3+scale_y_continuous(breaks=c(-20,0,20,40,60))&theme_bw()

#plot the pointiwse intervals only

ggplot()+
geom_ribbon(aes(ymin = lwrP, ymax = uprP,x=Day,group=Group,fill='red'),
            alpha = 0.1,
            show.legend = FALSE,
            data=pred) +
    geom_ribbon(aes( x=Day,ymin=(fit - 2*se.fit), 
                   ymax=(fit + 2*se.fit),
                   group=Group
                   ),
              alpha=0.1,
              fill='green',
              data=gam_predict,
            show.legend=FALSE,
            inherit.aes = FALSE)+
    geom_ribbon(data = ci, mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth,fill = 'yellow'),
inherit.aes = FALSE, alpha = 0.8,
show.legend=FALSE)

ggplot()+
    geom_ribbon(aes(ymin = lwrS, ymax = uprS,x=Day,group=Group),
                fill='yellow',
                alpha = 0.4,
                show.legend = FALSE,data=pred) +
geom_ribbon(data = si,
mapping = aes(ymin = lower, ymax = upper, x = Day, group = smooth),
fill ='red',
inherit.aes = FALSE, alpha = 0.4,
show.legend=FALSE)

```

From the plots, this is apparent:

The limits of the CI interval are the same in G.Simpson's code and in our original approach to the matter, they overlap perfectly. From our code however, the pointwise CI is slightly wider.

For the simultaneous CIs, the CI computed by G. Simpson's code is wider than the one we obtain using the `confint` function. Why is this?

```{r,poitwise-diff}
##this function determines the pointwise confidence interval of a difference between two smooths

difference_pointwise <- function(f1, f2, smooth, by_var, smooth_var, data, Xp, V, coefs, nrep = 1000) {
    ## make sure f1 and f2 are characters
    f1 <-  as.character(f1)
    f2 <-  as.character(f2)
    cnames <- colnames(Xp)
    ## columns of Xp associated with pair of smooths
    c1 <- grepl(gratia:::mgcv_by_smooth_labels(smooth, by_var, f1), cnames, fixed = TRUE)
    c2 <- grepl(gratia:::mgcv_by_smooth_labels(smooth, by_var, f2), cnames, fixed = TRUE)
    ## rows of Xp associated with pair of smooths
    r1 <- data[[by_var]] == f1
    r2 <- data[[by_var]] == f2

    ## difference rows of Xp for pair of smooths
    X <- Xp[r1, ] - Xp[r2, ]

    ######IMPORTANT: uncommenting the following two lines
    #removes the group means from the comparison######

    ## zero the cols related to other splines
    # X[, ! (c1 | c2)] <- 0

    ## zero out the parametric cols
    #X[, !grepl('^s\\(', cnames)] <- 0

    ## compute difference
    sm_diff <- drop(X %*% coefs)
    se <- sqrt(rowSums((X %*% V) * X))
    nr <- NROW(X)

    ## Calculate posterior simulation for smooths
    coefs_sim <- t(rmvn(nrep, rep(0, nrow(V)), V))
    rownames(coefs_sim) <- rownames(V)
    simDev <- X %*% coefs_sim
    absDev <- abs(sweep(simDev, 1, se, FUN = "/"))
    masd <- apply(absDev, 2, max)
    crit_s <- quantile(masd, prob = 0.95, type = 8)

    out <- list(smooth = rep(smooth, nr), by = rep(by_var, nr),
                level_1 = rep(f1, nr),
                level_2 = rep(f2, nr),
                diff = sm_diff, se = se,
                lower_s = sm_diff - crit_s * se,
                upper_s = sm_diff + crit_s*se)

    out <- new_tibble(out, nrow = NROW(X), class = "difference_smooth")
    ## Only need rows associated with one of the levels
    out <- bind_cols(out, data[r1, smooth_var])

    out
}

```

And this:

```{r,difference-smooths}
#this function calculates the pointwise (by calling pointwise_comparisons.R) CI and the simultaneous CI
# for a pairwise comparison between two smooths
difference_smooths <- function(model,
                               smooth,
                               n = 100,
                               ci_level = 0.95,
                               newdata = NULL,
                               partial_match = TRUE,
                               unconditional = FALSE,
                               frequentist = FALSE,
                               nrep = 10000,
                               include_means = TRUE,
                               ...) {
    if (missing(smooth)) {
        stop("Must specify a smooth to difference via 'smooth'.")
    }

    # smooths in model
    S <- gratia::smooths(model) # vector of smooth labels - "s(x)"
    # select smooths
    select <-
        gratia:::check_user_select_smooths(smooths = S, select = smooth,
                                           partial_match = partial_match)#,
    # model_name = expr_label(substitute(object)))
    sm_ids <- which(select)
    smooths <- gratia::get_smooths_by_id(model, sm_ids)
    sm_data <- map(sm_ids, gratia:::smooth_data,
                   model = model, n = n, include_all = TRUE)
    sm_data <- bind_rows(sm_data)
    by_var <- by_variable(smooths[[1L]])
    smooth_var <- gratia:::smooth_variable(smooths[[1L]])
    pairs <- as_tibble(as.data.frame(t(combn(levels(sm_data[[by_var]]), 2)),
                                     stringsAsFactor = FALSE))
    names(pairs) <- paste0("f", 1:2)

    Xp <- predict(model, newdata = sm_data, type = "lpmatrix")
    V <- gratia:::get_vcov(model, unconditional = unconditional,
                           frequentist = frequentist)
    coefs <- coef(model)

    out <- pmap(pairs, difference_pointwise, smooth = smooth, by_var = by_var,
                smooth_var = smooth_var, data = sm_data, Xp = Xp, V = V,
                coefs = coefs, nrep = nrep)
    out <- bind_rows(out)
    crit <- qnorm((1 - ci_level) / 2, lower.tail = FALSE)

    out <- add_column(out,
                      lower = out$diff - (crit * out$se),
                      upper = out$diff + (crit * out$se),
                      .after = 6L)
    out
}
```

Let's get the fit then:

```{r,testing}
diff_testing <- difference_smooths(gam_02, smooth = "s(Day)", newdata = newdat,
                     unconditional = TRUE, frequentist = FALSE,
                     n=100, partial_match = TRUE, nrep=10000)

```